\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother


\date{}

\begin{document}

\#MCBG 2032

\hypertarget{section}{%
\subsection{13/02/2108}\label{section}}

\hypertarget{phenylketonuria}{%
\subsubsection{Phenylketonuria}\label{phenylketonuria}}

single mutation can affect brain development however if the correct diet
is followed then all detrimental health consequences can be ignored.

\hypertarget{diabetes-type-ii}{%
\subsubsection{diabetes type II}\label{diabetes-type-ii}}

lifestyle disease, the genetic component is estimated to change the
likelihood of developing the disease by 2\%, but the main causative
factors of the disease relate to diet.

\hypertarget{lactose-intolerance.}{%
\subsubsection{Lactose intolerance.}\label{lactose-intolerance.}}

Individuals with lactose intolerance cant digest lactose, however the
bacteria in their gut van which leads to the build up of gas. this can
be very painful for the person in question.In humans lactose intolerance
would naturally occur after weening however as cow milk is a staple part
of most human diabetes the genes controlling lactase production are
never switched off and the ability to digest lactose is not normally
lost.

\hypertarget{himalayan-phenotype.}{%
\subsubsection{Himalayan Phenotype.}\label{himalayan-phenotype.}}

\hypertarget{cancer}{%
\subsubsection{Cancer}\label{cancer}}

\hypertarget{basics}{%
\paragraph{basics}\label{basics}}

A cancer is a group of cells whose proliferation is uncontrolled, and
which can spread to other locations in the body which are normally
populated bu other cell types.

\hypertarget{benigin-tumours}{%
\subparagraph{benigin tumours}\label{benigin-tumours}}

grow but do not spread.

\hypertarget{malignant-tumours}{%
\subparagraph{malignant tumours}\label{malignant-tumours}}

over proliferate and invade other body tissues/ areas in the body.

\hypertarget{types-of-cancers}{%
\paragraph{types of cancers}\label{types-of-cancers}}

\hypertarget{carcinomas}{%
\subparagraph{Carcinomas}\label{carcinomas}}

derived form epithelial cells

\hypertarget{sarcomas}{%
\subparagraph{Sarcomas}\label{sarcomas}}

derived from connective or muscle tissues. Known as Osteosarcoma in the
case of bone tissue and Kaposi sarcoma in the case of soft connective
tissue.

\hypertarget{lymphoma-and-leukaemias}{%
\subparagraph{Lymphoma and leukaemias}\label{lymphoma-and-leukaemias}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Cancers of the hematopoietic system.
\item
  Lymphoma (solid tumour)
\item
  Leukaemia excess of circulating immature blood cell precursors.
  \#\#\#\#\# Cancer of the nervous system brain and central nervous
  system.
\end{enumerate}

NOTE: most common in children

\hypertarget{germ-line-cancer.}{%
\subparagraph{Germ-line cancer.}\label{germ-line-cancer.}}

\hypertarget{cancers-of-breast-prostate-lung-pancreas-and-colon.}{%
\subparagraph{Cancers of breast, prostate, lung, pancreas, and
colon.}\label{cancers-of-breast-prostate-lung-pancreas-and-colon.}}

\hypertarget{causes-of-cancer.}{%
\subsubsection{Causes of cancer.}\label{causes-of-cancer.}}

In most cases cancer is not inherited. In breast cancer/ovarian cancer
BS1, BS2 play some role in genetic disposition, perhaps 5-10\%

\hypertarget{hereditary.}{%
\subsubsection{hereditary.}\label{hereditary.}}

Cancer normally sets in during old age as cells must accumulate a series
of a specific set of notations, (exactly which set of mutations
accumulated lead to cancer is highly individual)

\hypertarget{environmental-factors}{%
\paragraph{Environmental factors}\label{environmental-factors}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  UV
\item
  X rays
\item
  Alcohol
\item
  Overcooked food (hetercyclic amines,polycyclic aromatic
  hydrocarbons-meat. acrylamide- potatoes
\item
  Azo dyes
\item
  tartrazine (food colourants)
\item
  Nitrate cured foods
\item
  Pesticides
\end{enumerate}

\hypertarget{cancer-initiators}{%
\paragraph{Cancer initiators}\label{cancer-initiators}}

agents which cause DNA damage (mutagens). these factors may be chemical,
biological (such as HPV- human papilloma virus cause latent genetic
damage- Cancer predisposition.

\hypertarget{cancer-promoters}{%
\paragraph{Cancer Promoters}\label{cancer-promoters}}

Promotes excessive proliferation (does not directly damage DNA).

\hypertarget{examples}{%
\subparagraph{Examples}\label{examples}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  wounding
\item
  phorbol esters
\item
  HRT/ oestrogen (breast cancer)
\item
  hepatitis B (promotes stomach cancer)
\item
  HIV (Kaposi sarcoma, this disease is always present but only manifests
  itself in immunocompromised individuals)
\end{enumerate}

\hypertarget{stages-of-cancer-development}{%
\paragraph{stages of cancer
development}\label{stages-of-cancer-development}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  normal cells (mutation occcurs)
\item
  increased ability to proliferate (mutation)
\item
  1-2mm tumour of rapidly proliferating cells which do not undergo
  apoptosis of differentiation. (mutation)
\item
  vascularized growing tumour (mutation)
\item
  large tumour capable of invading near by tissue (metastasis)
\end{enumerate}

NOTE: cancer results from a series of somatic mutations, affecting the
same cell, and different cancers are genetically heterogeneous.

\hypertarget{genetic-instability-in-cancers.}{%
\paragraph{Genetic instability in
cancers.}\label{genetic-instability-in-cancers.}}

deficient local DNA repair leads to the accumulation of point mutations
increased chromosomal instability and gross genome abnormalities.

\hypertarget{summary-of-key-properties}{%
\paragraph{summary of key properties}\label{summary-of-key-properties}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  disregards ex/in growth regulation signals.
\item
  avoid apoptosis, differentiation and replicative senescence
\item
  genetically unstable.
\item
  invasive
\item
  metastatic (survive and proliferate in foreign) sites.
\end{enumerate}

\hypertarget{penetrance}{%
\paragraph{Penetrance}\label{penetrance}}

percentage of the population who demonstrate at least some degree fo
phenotypic expression.

\hypertarget{expressivity}{%
\paragraph{Expressivity}\label{expressivity}}

reflects the range of expressions of the gene/allele present in the
population.

\hypertarget{genetic-complementation}{%
\paragraph{Genetic complementation}\label{genetic-complementation}}

two different mutations in heterozygous condition affecting the same
protein/ pathway can compliment each other to cause a novel phenotypic
effect.

\hypertarget{examples-1}{%
\subparagraph{examples}\label{examples-1}}

rare form of albinism: both mother and father were normally pigmented
and there families had minimal instances of albinism. however both
children where albino' s . Analysis found that this albinism stemmed not
from those genes and mutation most commonly associated with albinism but
rather from a combination of recessives (heterozygous) from both father
and mother (mutations in TYRP1 gene), which on their own would have
little or no effect on the pigment production pathway, but when combined
reduced pigment production significantly.

NOTE: Rufus albinism leads to a phenotype with reddish hair, lighter
skin, and blue grey eyes.

\hypertarget{forked-line-probability-method.}{%
\paragraph{Forked line probability
method.}\label{forked-line-probability-method.}}

\#\#chromosome level (revise mitosis and meosis) copy slides chromosome
basic structure \#\#\#\# chromosome groups

\hypertarget{meta-centric}{%
\subparagraph{meta-centric}\label{meta-centric}}

centromere is half way up the length of the chromosome

\hypertarget{submetacentric}{%
\subparagraph{submetacentric}\label{submetacentric}}

centromere is more to the one side of the chromosome than the other,

\hypertarget{acrocentric}{%
\subparagraph{Acrocentric}\label{acrocentric}}

the centromere is very far to the one side of the chromosome, with a
long arm containing most of the genes and a short arm containing
predominantly temolmeric DNA.

\hypertarget{telocentric}{%
\subparagraph{Telocentric}\label{telocentric}}

humans do not posses any telocentric chromosomes, but certain insect of
crustation species do.

\hypertarget{holocentric}{%
\subparagraph{Holocentric}\label{holocentric}}

Centromere like structure exist along the entire length of the
chromosome. this may decrease the chances of faulty division/
segregation.

\hypertarget{chromosome-banding}{%
\subparagraph{chromosome banding}\label{chromosome-banding}}

bands were named and used to locate specific genes. the banding patterns
are due to uneven DNA densities in the coiled structure of the
chromosome.

\hypertarget{size-and-shape-of-different-chromosomes.}{%
\subparagraph{Size and shape of different
chromosomes.}\label{size-and-shape-of-different-chromosomes.}}

\hypertarget{chromosome-level-mutations}{%
\subsubsection{Chromosome level
mutations}\label{chromosome-level-mutations}}

\hypertarget{aneuploidy-spelling}{%
\paragraph{Aneuploidy (Spelling)}\label{aneuploidy-spelling}}

Each cell is has at least one extra chromosome or is missing at least
one chromosome

\hypertarget{trisomy}{%
\subparagraph{Trisomy}\label{trisomy}}

when an individual inherits three copies of a particular chromosome. In
Humans only three of all the possible trisomies are viable, (as in
individuals with these mutations will still be born and not terminated
during pregnancy)

of these three (Trisomy 13, 18, and 21) only individuals with trisomy 21
can survive past the first few years of childhood. Individuals with
Trisomy 21 have down syndrome.

Down syndrome (Trisomy 21)

Frequency: 1/1000 Effects 1. mental retardation 2. short stature 3.
heart disease 4. shortened life span.

Edward Syndrome (Trisomy 18)

Effects: 1. kidney and intestinal malformation 3. heart defects. 4.
mental retardation

NOTE: only 8\% survival past the first year.

Patua syndrome (Trisomy 13 )

Effects: 1. Kidney and hear defects 2. polidactily (too many digits) 3.
Nervous system abnormalities 4. Death within the 1st year.

NOTE: the smaller relative size of chromosome 21, meaning that it
contains less important genes may be related to its increased viability
.

\hypertarget{monosomy}{%
\subparagraph{Monosomy}\label{monosomy}}

when an individual inherits only one copy of a particular chromosome.

\hypertarget{nullisomy}{%
\subparagraph{Nullisomy}\label{nullisomy}}

having no copies of a particular chromosome.

\hypertarget{polyploidy}{%
\paragraph{Polyploidy}\label{polyploidy}}

\hypertarget{structural-rearrangements}{%
\paragraph{Structural rearrangements}\label{structural-rearrangements}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion
\item
  duplication
\item
  translocation
\item
  inversion
\item
  fission
\item
  fusion
\end{enumerate}

\hypertarget{non-disjuncture}{%
\paragraph{non disjuncture}\label{non-disjuncture}}

(review mitosis and meiosis) If non disjuncture happens in meiosis it is
generally worse than if it happens in mitosis as it will affect the
whole organism.

\hypertarget{down-syndrome}{%
\subparagraph{Down Syndrome}\label{down-syndrome}}

occurs in about 1 of a thousand live births. originates from non
disjunction in the egg cell rather than the sperm cell because the
arrested development of the egg cells leads to decay of the separating
and marking proteins as well as the spindle fibre leaving the whole
process open to more error. 8\% of individuals survive for one year ?

\hypertarget{turner-syndrome-monosomy-x}{%
\subparagraph{turner Syndrome monosomy
X}\label{turner-syndrome-monosomy-x}}

Frequency 1/20 000 (only occurs in females ) Effects: 1. webbed neck 2.
short stature 3. underdeveloped ovaries and lack of secondary sex
characteristics.

\hypertarget{klinefelter-syndrome-xxy}{%
\subparagraph{Klinefelter syndrome XXY}\label{klinefelter-syndrome-xxy}}

Too many X's ,XXY , XXXY, XXXXY, XXXXXY. (XXYY) Effects: 1. Decrease in
testosterone levels 2. the more X's the higher the chance of brain
damage.

\hypertarget{xyy}{%
\subparagraph{XYY}\label{xyy}}

XYY has the phenotype of a normal make. men with 2 Y chromosomes tend to
be tall as there is a cumulative effect adding to height. (they may also
exhibit increased aggression but this in unclear)

\hypertarget{xxx}{%
\subparagraph{XXX}\label{xxx}}

Frequency: 1/1000 Effects: 1. Normal female with normal fertility.

NOTE: this conditions is seldom diagnosed.

\hypertarget{polyploidy-1}{%
\paragraph{Polyploidy}\label{polyploidy-1}}

Three or more complete sets of chromosomes present in somatic cells.
\#\#\#\# 2n-20n caused by non reduced gamete fomration, which is very
rare in males and non viable.

most often flowering plants are polyploid, they are even specifically
breed to have more chromosome sets as this usually increases fruit and
flower size.

\hypertarget{examples-2}{%
\subparagraph{Examples}\label{examples-2}}

water melon must be bread from a tetraploid(?) and a diploid to get a
tripliod infertile plant.

Kiwi fruit 12-16 copies

Strawberries 4 copies (tetraploid)

Frogs \emph{Xenopis levis} 4n \emph{Xenopus tropicalis} 2n

NOTE: plants are better adapted to polyploid because: 1. they are not as
confined to a set physical form so different in growth and development
genes operation levels are not so important 2. they can reproduce
vegetatively so polyploid individuals aren't as severely evolutionarily
disadvantaged. 3. less precise sperm targeting is necessary.

\hypertarget{deletions}{%
\paragraph{Deletions}\label{deletions}}

deletions often occur by chromosome breakage during the cell cycle.
These deletions may be terminal of interstitial. If the centromere is
lost, the entire chromosome will be lost.

\hypertarget{piece-of-chrome}{%
\subparagraph{Piece of Chrome}\label{piece-of-chrome}}

sometimes multiple genes are lost when both strands break and a part of
the chromosome is permanently lost. A specific deletion on chromosome 5.
(where the entire p arm is deleted leads to a serious syndrome)

\hypertarget{cri-du-chat-5p-}{%
\subparagraph{Cri du chat (5p-)}\label{cri-du-chat-5p-}}

Frequency: 1/20000-50000 Effects: 1. Intellectual disability, 2. low
muscle tone 3. microcephaly 4. distinctive facial features. 5. normal
life expectancy

\hypertarget{duplications}{%
\paragraph{Duplications}\label{duplications}}

Cause: retro transposition, or non-allelic homologous recombination in
repeat rich areas.

A small region of a chromosome is repeated, and placed next to the
existing copy (in tandem). duplications tend to happen in high
repetitive sequences of DNA. These repeated regions mean that the
chromosome will note line up very well with its homologous pair.

passed on by unequal order, and may or may not have a phenotypic effect.

\hypertarget{huntington-purea}{%
\paragraph{Huntington Purea}\label{huntington-purea}}

Repeats alter gene function, which can lead to brain damage. Repeats may
have a positive effect because they allow one copy of a gene to evolve
independently, specialise in a different function while the original
function is still conserved by the other copy of the gene.

\hypertarget{inversions.}{%
\paragraph{Inversions.}\label{inversions.}}

Cause: non-allelic homologous recombination Effects: 1. pairing in
meiosis occurs via looping 2. cross over in the looped region can cause
major deletions or duplications.

\hypertarget{translocation}{%
\paragraph{Translocation}\label{translocation}}

Effects: 1. novel chromosomes are generated 2. individuals are usually
infertile. 3. can result in aneuploidy (such as 14q21q inherited down
syndrome) 4.

\hypertarget{terminal}{%
\subparagraph{terminal}\label{terminal}}

A piece breaks off a chromosome and sticks to the end of a
non-homologous chromosome.

\hypertarget{reciprocal}{%
\subparagraph{reciprocal}\label{reciprocal}}

Exchange of pieces between non-homologous chromosomes \#\#\#\#
Robertsonian

two Acrocentric chromosomes lose their short arms and get stuck
together.

\hypertarget{fission}{%
\paragraph{Fission}\label{fission}}

deletion (loss of acentric piece)'

\hypertarget{fusion}{%
\paragraph{Fusion}\label{fusion}}

bicentric chromosomes (breakage in mitosis)

\hypertarget{breakage-fusion-bridge-cycles-in-cancer.}{%
\subparagraph{breakage fusion bridge cycles in
cancer.}\label{breakage-fusion-bridge-cycles-in-cancer.}}

Copy image. 1. end of a chromosome breaks off (during meiosis) 2. one
daughter cell inherits the chromosome lacking a telomere. 3. the new
cell enters S phase and replicates its DNA (it now has two sticky DNA
ends right next to each other. ) 4. loose DNA strands bind to form
continuous DNA loop. 5. fused sister chromatids are pulled apart during
mitosis, causing breakage at a new sight 6. one daughter cell inherits
chromosome with duplicated genes and again a free sticky end. this cycle
can then begin again leading to even further chromosome distortion.

\hypertarget{imprinting}{%
\paragraph{Imprinting}\label{imprinting}}

\hypertarget{hinny-vs-mule}{%
\subparagraph{Hinny vs mule}\label{hinny-vs-mule}}

Mule

horse mother donkey father hardy and obedient

Hinny

donkey mother, horse father. temperamental, untrainable.

\hypertarget{intersistial-deletion-of-chromosome-15}{%
\subparagraph{Intersistial deletion of chromosome
15}\label{intersistial-deletion-of-chromosome-15}}

Prader -Willi syndrome

Cause: deletion from father, ZNF127 and IPW inactivated on the maternal
copy.\\
effects: 1. mental retardation 2. obesity 3. diminished growth.

Angelman syndrome

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion from mother: UBE3A inactivated on paternal copy.\\
\item
  mental retardation
\item
  epilepsy
\item
  lack of motor development.
\end{enumerate}

\hypertarget{x-inactivation}{%
\paragraph{X inactivation}\label{x-inactivation}}

the inactivation of one of two X chromosomes in females via DNA
methylation.

NOTE: which X chromosome is inactivated varies randomly from cell to
cell.this inactivation process turns the inactive X into a bar body,
(highly methylated and rolled up in histone proteins).

\hypertarget{mcbg-2036}{%
\section{MCBG 2036}\label{mcbg-2036}}

\hypertarget{section-1}{%
\subsection{14/02/2018}\label{section-1}}

\hypertarget{acids-and-bases.}{%
\subsubsection{acids and bases.}\label{acids-and-bases.}}

\hypertarget{ph}{%
\paragraph{pH}\label{ph}}

\(pH=-log_{10}[H^(+)]\)

\#\#\#\#ka

\(k_{a}=\frac{[A^{-}][H^+]}{[HA]}\)

\hypertarget{pka}{%
\paragraph{pKa}\label{pka}}

\(pka=\log_{10}(pk_a)\)

\hypertarget{dependant-and-independant-variables.}{%
\paragraph{dependant and independant
variables.}\label{dependant-and-independant-variables.}}

\hypertarget{buffers}{%
\paragraph{Buffers}\label{buffers}}

NOTE: buffers may be temperature sensitive, for example Tris buffers.
phosphate buffers are more reliable but it is still best to take
temperature into consideration

\hypertarget{indicator-dyes.}{%
\paragraph{Indicator dyes.}\label{indicator-dyes.}}

\hypertarget{examples-3}{%
\subparagraph{examples}\label{examples-3}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  phenlythyelian (spelling)
\item
  bromothymyl blue (spelling)
\item
  methyl orange
\end{enumerate}

\hypertarget{paper-strips}{%
\subparagraph{paper strips}\label{paper-strips}}

inaccurate but useful for rough estimations of pH or to identify changes
in pH

\hypertarget{electrode}{%
\subparagraph{Electrode}\label{electrode}}

accurate to two decimal places.

\#APES

\hypertarget{survivor-ship-curves}{%
\subsection{Survivor ship curves}\label{survivor-ship-curves}}

insert appropriate graphs.

\hypertarget{type1-1}{%
\subsubsection{Type1 1}\label{type1-1}}

\hypertarget{typical-populations}{%
\paragraph{Typical populations}\label{typical-populations}}

domestic pets and humans in the developed world

\hypertarget{type-2}{%
\subsubsection{Type 2}\label{type-2}}

\hypertarget{typical-populations-1}{%
\paragraph{Typical populations}\label{typical-populations-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  seeds in a seed bank
\item
  marmites.
\end{enumerate}

\hypertarget{type-3}{%
\subsubsection{Type 3}\label{type-3}}

\hypertarget{typical-populations-2}{%
\paragraph{typical populations}\label{typical-populations-2}}

large mammals.

\hypertarget{reproductive-rate}{%
\subsubsection{Reproductive Rate}\label{reproductive-rate}}

combination of many factors. 1. length of reproductive season. 2.
(reproduced output ) number of offspring produced. 3. Investment 4.
generation length (time between birth of mother and birth of daughter)
5. age of maturity.

\hypertarget{potential-population-rate-of-change.}{%
\paragraph{potential population rate of
change.}\label{potential-population-rate-of-change.}}

how fast populations can change. \(r_{m} ~or~ r\) are common symbols for
the maximum growth rate.

\hypertarget{k-selecting}{%
\subsubsection{K selecting}\label{k-selecting}}

adaptions which favour ability to survive over the ability to reproduce.
well adapted to predictable, crowded environments with significant
resource scarcity.

\hypertarget{general-traits}{%
\paragraph{general traits}\label{general-traits}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  large
\item
  can deffer reproduction to a later time.
\item
  iteroparous.
\item
  high investment per offspring
\item
  body resources invested more into survival than reproduction.
\end{enumerate}

\hypertarget{r-selecting}{%
\subsubsection{R selecting}\label{r-selecting}}

unpredictable environments, periods/ areas of sudden abundance.

\hypertarget{general-traits-1}{%
\paragraph{general traits}\label{general-traits-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  small(er)
\item
  earlier maturity
\item
  invest in reproduction versus survival
\item
  many offspring but with low investment per offspring.
\end{enumerate}

\hypertarget{variation-in-abundance-over-time.}{%
\paragraph{Variation in abundance over
time.}\label{variation-in-abundance-over-time.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  patchy resource allocation
\item
  inter-species interactions (predation, symbiosis)
\item
  environment
\item
  social structure
\end{enumerate}

\hypertarget{resources.}{%
\subparagraph{resources.}\label{resources.}}

physicochemical. conditions.\\
1. temperature 2. humidity 3. occupancy.

\hypertarget{defining-environments.}{%
\subsection{defining environments.}\label{defining-environments.}}

environments are a smooth continuum of suitable (good) to unsuitable
(bad). There is a threshold below which conditions are too ad to sustain
a particular organism, below this line there are no available habitats.
above this line there is spectrum of habitats varying from poor to good,
with associated increase in survival and reproductive rates. \#\#\#\#\#
Examples Elephants in the Kruger

because of artificial water easy for herds with young elephants to move
around which otherwise they would be unable to migrate. artificial water
sources. this artificially increased the carrying capacity.

\#CHEM 2003

\hypertarget{valence-bond-theory}{%
\subsection{Valence Bond Theory}\label{valence-bond-theory}}

\hypertarget{history}{%
\subsubsection{History}\label{history}}

Valence bond theory was the First quantum mechanical theory developed.
It was developed by Walter Heitler, friz London and Linus Pauling, and
helped to explain concepts which were poorly predicted by VSPER theory
such as diamagnetic/ paramagnetic properties.

\hypertarget{basics-1}{%
\subsubsection{Basics}\label{basics-1}}

\hypertarget{wave-functions.}{%
\paragraph{Wave functions.}\label{wave-functions.}}

the wave function of an electron pair is formed by superimposing the
wave function for the separate fragments of a molecule.

\hypertarget{electrons}{%
\subparagraph{Electrons}\label{electrons}}

Each electron will have a wave function \(\psi\) that describes the
behaviour of that electron. (section 1.2 in book)

the probability of finding an electron at any given location is given by
\(\chi= \psi^{2}\)

\(/chi\) is used to represent the orbital of a given electron.

\hypertarget{bonds}{%
\subparagraph{bonds}\label{bonds}}

each bond will also have a wave function which describes the movement of
electrons within that orbital.

examples

the bond between two hydrogen can be expressed as
\(\psi= \chi_{A}(1)\chi_{B}(2)\) ( or \(\psi=\chi_{A}(2)\chi_{B}(1)\) is
just as valid )

\end{document}
