---
title: "Concepts Notes"
output: html_notebook
---


# Genetics

## Basics

### Course outline
1) extension of Mendelian genetics
2) Chromosomes structure and changes
3) Vertebrate development

### Definitions

## Interactions between Genes, Alleles, and Phenotypes

### Background

#### Penetrance
 percentage of the population who demonstrate at least some degree fo phenotypic expression.

#### Expressivity
reflects the range of expressions of the gene/allele present in the population.  


#### Genes 
A gene is a collection/ group of alleles which could occur the same locus and
which are associated with the same traits.

### Lethal alleles
alleles which are incompatible with the organisms continued survival

#### dominance
most lethal alleles are recessive because dominant lethal alleles produce non
functioning proteins only so any individual with even one of the dominant allele
would be unlikely to survive to reproductive age, and so the Allele would very
quickly be bread out of the population.

### Incomplete dominance

#### Genetic/ molecular cause
The recessive allele does not produce any functioning protein.
The dominant allele does function protein but not in high enough quantities to
result in the same phenotypic effect as a heterozygous dominant

#### phenotypic effect
A heterozygous individual will display a phenotype intermediate to that of
homozygous dominant and homozygous recessive

### Dominance

#### Genetic/molecular cause
The recessive allele does not produce any functioning protein.
The dominant allele does function protein, which it produces in comparative
amount to a homozygous dominant individual

#### Phenotypic consequence.
the heterozygous phenotype is identical to that of the heterozygous dominant
individual.

### Co-dominance

#### Genetic/ molecular cause
Both alleles are transcribed/translated to form a functioning protein

#### phenotypic effect
The effect comes from the full expression of both alleles

### Dominance hierarchies.
It is often the case that there will be a number of alleles for one gene. In
which case a dominance hierarchy is usually present such that allele A is
Dominant to B is dominant to C and so forth (ie dominance in a hierarchy is
 transitive)

### Pleiotropy 
Genetic/molecular cause
One gene locus affects the expression of several characteristics.
Phenotypic cause
Phenotype of one characteristic will change in conjunction with the phenotype of
another characteristic

### Important examples
BG 2032



#### Notes about colour
As   many genetic examples are based on colour it is worth understanding the basics
of colouring in animals.
colour results from pigments which are proteins, individual genes can either code
directly for the production of these proteins or as an intermediate/enzyme
which aids in the production of these proteins.

##### White
White does not result from a specific pigment but rather from a lack of pigment.

##### Co-dominance and Incomplete dominance
A co-dominant colour combination can look like an intermediate if it is the
combination of two pigments, however in the case where one of the phenotypes is
white then a intermediate can only be a result of incomplete dominance

### Sex linked traits

#### Sex limited traits,



## Genetic basis of disease. 

### Phenylketonuria
 single mutation
can affect brain development however if the correct diet is followed then all detrimental health consequences can be ignored.


### diabetes type II
lifestyle disease, the genetic component is estimated to change the likelihood of developing the disease by 2%, but the main causative factors of the disease relate to diet.




### Lactose intolerance.
Individuals with lactose intolerance cant digest lactose, however the bacteria in their gut van which leads to the build up of gas. this can be very painful for the person in question.In humans lactose intolerance would naturally occur after weening however as cow milk is a staple part of most human diabetes the genes controlling lactase production are never switched off and the ability to digest lactose is not normally lost.


### Cancer
 
### Hereditary.
Cancer normally sets in during old age as cells must accumulate a series of a specific set of notations, (exactly which set of mutations accumulated lead to cancer is highly individual)

#### Environmental factors
1. UV
2. X rays
3. Alcohol
4. Overcooked food (hetercyclic amines,polycyclic aromatic hydrocarbons-meat. acrylamide- potatoes
5. Azo dyes
6. tartrazine (food colourants)
7. Nitrate cured foods
8. Pesticides

#### Cancer initiators
agents which cause DNA damage (mutagens). these factors may be chemical, biological (such as HPV- human papilloma virus cause latent genetic damage- Cancer predisposition.

#### Cancer Promoters

Promotes excessive proliferation (does not directly damage DNA).

##### Examples
1. wounding
2. phorbol esters
3. HRT/ oestrogen (breast cancer)
4. hepatitis B (promotes stomach cancer)
5. HIV (Kaposi sarcoma, this disease is always present but only manifests itself in immunocompromised individuals)

#### stages of cancer development  
1. normal cells (mutation occcurs)
2. increased ability to proliferate (mutation)
3. 1-2mm tumour of rapidly proliferating cells which do not undergo apoptosis of differentiation. (mutation)
4. vascularized growing tumour (mutation)
5. large tumour capable of invading near by tissue (metastasis)

NOTE: cancer results from a series of somatic mutations, affecting the same cell, and different cancers are genetically heterogeneous.

#### Genetic instability in cancers.
deficient local DNA repair leads to the accumulation of point mutations increased chromosomal instability and gross genome abnormalities.

#### summary of key properties
1. disregards ex/in growth regulation signals.
2. avoid apoptosis, differentiation and replicative senescence
3. genetically unstable.
4. invasive
5. metastatic (survive and proliferate in foreign) sites.


## Genetic complementation
two different mutations in heterozygous condition affecting the same protein/ pathway can compliment each other to cause a novel phenotypic effect.

##### examples
rare form of albinism: both mother and father were normally pigmented and there families had minimal instances of albinism. however both children where albino' s . Analysis found that this albinism stemmed not from those genes and mutation most commonly associated with albinism but rather from a combination of recessives (heterozygous) from both father and mother (mutations in TYRP1 gene), which on their own would have little or no effect on the pigment production pathway, but when combined reduced pigment production significantly.

NOTE: Rufus albinism leads to a phenotype with reddish hair, lighter skin, and blue grey eyes.

## Forked line probability method.
 
#### basics
A cancer is a group of cells whose proliferation is uncontrolled, and which can spread to other locations in the body which are normally populated bu other cell types.

##### benigin tumours
grow but do not spread.

##### malignant tumours
over proliferate and invade other body tissues/ areas in the body.  
 
#### types of cancers
 
##### Carcinomas
 derived form epithelial cells
 
##### Sarcomas
 derived from connective or muscle tissues. Known as Osteosarcoma in the case of bone tissue and Kaposi sarcoma in the case of soft connective tissue.  
 
##### Lymphoma and leukaemias
1 . Cancers of the hematopoietic system.
2. Lymphoma (solid tumour)
3. Leukaemia excess of circulating immature blood cell precursors.
##### Cancer of the nervous system
brain and central nervous system.

NOTE: most common in children
 
##### Germ-line cancer.

##### Cancers of breast, prostate, lung, pancreas, and colon.

#### Causes of cancer.
I n most cases cancer is not inherited. In  breast cancer/ovarian cancer BS1, BS2 play some role in genetic disposition, perhaps 5-10%
 

## Chromosome level genetics
(re vise mitosis and meosis)
copy slides

### Background 

#### Chromosome basic structure

#### Chromosome groups

##### meta-centric
centromere is half way up the length of the chromosome

##### submetacentric
centromere is more to the one side of the chromosome than the other,

##### Acrocentric
the centromere is very far to the one side of the chromosome, with a long arm containing most of the genes and a short arm containing predominantly temolmeric DNA.

##### Telocentric  
humans do not posses any telocentric chromosomes, but certain insect of crustation species do.

##### Holocentric
Centromere like structure exist along the entire length of the chromosome. this may decrease the chances of faulty division/ segregation.

#### Chromosome banding
bands were named and used to locate specific genes. the banding patterns are due to uneven DNA densities in the coiled structure of the chromosome.


### Chromosome level mutations

#### Aneuploidy (Spelling)
Each cell is has at least one extra chromosome or is missing at least one chromosome

##### Trisomy
when an individual inherits three copies of a particular chromosome.
In Humans only three of all the possible trisomies are viable, (as in individuals with these mutations will still be born and not terminated during pregnancy)

of these three (Trisomy 13, 18, and 21) only individuals with trisomy 21 can survive past the first few years of childhood. Individuals with Trisomy 21 have down syndrome.

###### Down syndrome (Trisomy 21)
Frequency: 1/1000
Effects
1. mental retardation
2. short stature
3. heart disease
4. shortened life span.

###### Edward Syndrome (Trisomy 18)
Effects:
1. kidney and intestinal malformation
3. heart defects.
4. mental retardation

NOTE: only 8% survival past the first year.

###### Patua syndrome (Trisomy 13 )
Effects:
1. Kidney and hear defects
2. polidactily (too many digits)
3. Nervous system abnormalities
4. Death within the 1st year.



NOTE: the smaller relative size of chromosome 21, meaning that it contains less important genes may be related to its increased viability .

##### Monosomy
when an individual inherits only one copy of a particular chromosome.

##### Nullisomy
having no copies of a particular chromosome.

#### Polyploidy

#### Structural rearrangements
1. deletion
2. duplication
3. translocation
4. inversion
5. fission
6. fusion

#### non disjuncture
(review mitosis and meiosis)
If non disjuncture happens in meiosis it is generally worse than if it happens in mitosis as it will affect the whole organism.

##### Down Syndrome
occurs in about 1 of a thousand live births. originates from non disjunction in the egg cell rather than the sperm cell because the arrested development of the egg cells leads to decay of the separating and marking proteins as well as the spindle fibre leaving the whole process open to more error.
8% of individuals survive for one year ?

##### turner Syndrome monosomy X
Frequency 1/20 000 (only occurs in females )
Effects:
1. webbed neck
2. short stature
3. underdeveloped ovaries and lack of secondary sex characteristics.

##### Klinefelter syndrome XXY
Too many X's ,XXY , XXXY, XXXXY, XXXXXY. (XXYY)
Effects:
1. Decrease in testosterone levels
2. the more X's the higher the chance of brain damage.

##### XYY
XYY has the phenotype of a normal make.
men with 2 Y chromosomes tend to be tall as there is a cumulative effect adding to height. (they may also exhibit increased aggression but this in unclear)

##### XXX
Frequency: 1/1000
Effects:
1. Normal female with normal fertility.

NOTE: this conditions is seldom diagnosed.


#### Polyploidy
Three or more complete sets of chromosomes present in somatic cells.
#### 2n-20n
caused by non reduced gamete fomration, which is very rare in males and non viable.

most often flowering plants are polyploid, they are even specifically breed to have more chromosome sets as this usually increases fruit and flower size.


##### Examples

water melon
must be bread from a tetraploid(?) and a diploid to get a tripliod
infertile plant.

Kiwi fruit
12-16 copies

Strawberries
4 copies (tetraploid)

Frogs
_Xenopis levis_  4n
_Xenopus tropicalis_ 2n

NOTE: plants are better adapted to polyploid because:
1. they are not as confined to a set physical form so different in growth and development genes operation levels are not so important
2. they can reproduce vegetatively so polyploid individuals aren't as severely evolutionarily disadvantaged.
3. less precise sperm targeting is necessary.



#### Deletions
deletions often occur by chromosome breakage during the cell cycle. These deletions may be terminal of interstitial. If the centromere is lost, the
entire chromosome will be lost.

##### Piece of Chrome
sometimes multiple genes are lost when both strands break and a part of the chromosome is permanently lost. A specific deletion on chromosome 5. (where the entire p arm is deleted leads to a serious syndrome)

##### Cri du chat (5p-)
Frequency: 1/20000-50000
Effects:
1. Intellectual disability,
2. low muscle tone
3. microcephaly
4. distinctive facial features.
5. normal life expectancy  

#### Duplications
Cause: retro transposition, or non-allelic homologous recombination in repeat rich areas.

A small region of a chromosome is repeated, and placed next to the existing copy (in tandem). duplications tend to happen in high repetitive sequences of DNA. These repeated regions mean that the chromosome will note line up very well with its homologous pair.

passed on by unequal order, and may or may not have a phenotypic effect.


#### Huntington Purea
Repeats alter gene function, which can lead to brain damage.
Repeats may have a positive effect because they allow one copy of a gene to evolve independently, specialise in a different function while the original function is still conserved by the other copy of the gene.

#### Inversions.
Cause: non-allelic homologous recombination
Effects:
1.  pairing in meiosis occurs via looping
2. cross over in the looped region can cause major deletions or duplications.

#### Translocation
Effects:
1. novel chromosomes are generated
2. individuals are usually infertile.
3. can result in aneuploidy (such as 14q21q inherited down syndrome)
4.

##### terminal
A piece breaks off a chromosome and sticks to the end of a non-homologous chromosome.

##### reciprocal
Exchange of pieces between non-homologous chromosomes
#### Robertsonian

two Acrocentric chromosomes lose their short arms and get stuck together.

#### Fission
deletion (loss of acentric piece)'

#### Fusion
bicentric chromosomes (breakage in mitosis)

##### breakage fusion bridge cycles in cancer.
Copy image.
1. end of a chromosome breaks off (during meiosis)
2. one daughter cell inherits the chromosome lacking a telomere.
3. the new cell enters S phase and replicates its DNA (it now has two sticky DNA ends right next to each other.
)
4. loose DNA strands bind to form continuous DNA loop.
5. fused sister chromatids are pulled apart during mitosis, causing breakage at a new sight
6. one daughter cell inherits chromosome with duplicated genes and again a free sticky end. this cycle can then begin again leading to even further chromosome distortion.

#### Imprinting

##### Hinny vs mule

###### Mule
horse mother donkey father
hardy and obedient

###### Hinny
donkey mother, horse father.
temperamental, untrainable.

##### Intersistial deletion of chromosome 15

###### Prader -Willi syndrome
Cause: deletion from father, ZNF127 and IPW inactivated on the maternal copy.  
effects:
1. mental retardation
2. obesity
3. diminished growth.

###### Angelman syndrome
1. deletion from mother: UBE3A inactivated on paternal copy.  
2. mental retardation
2. epilepsy
3. lack of motor development.

#### X inactivation
the inactivation of one of two X chromosomes in females via DNA methylation.

NOTE: which X chromosome is inactivated varies randomly from cell to cell.this inactivation process turns the inactive X into a bar body, (highly methylated and rolled up in histone proteins).

#### Frog colour

##### details
the skin colour of a certain from is controlled by one allele. There are three
different phenotypes observed, these are red, blue, and purple.

##### explanation
This is an example of co-dominance as the purple colour comes from a combination
of the red and blue pigments. that is both red and blue pigments have been
produced. We can conclude that the purple from is heterozygous, the blue frog
is homozygous blue and the red frog is homozygous red.

#### Fruit flies eye colour.

##### details.
the colour of fruit flies eyes is determined by two gene loci. a fly with
dominant alleles at the red locus will be red eyed, while a fly with dominant
 alleles at the brown locus will be brown eyed. However a fly with both
dominant brown and cinnabar will be white eyed.

##### explanation
the white eyed phenotype is due to a form of inhibition similar to epistasis
stops either of the pigments form being produced if both pigment genes are
present.

#### Rabbit Skin Colour

##### details
the skin colour of rabbits is determined by one gene locus, which has four
possible Alleles.
All alleles code for the enzyme tyrosinase which acts to convert tyrosine into
dopamine a critical first set in the production of hair pigment.

###### #1 non functional
The most recessive Allele codes for a nonfunctioning version of tyrosinase
leading to albino rabbit.

###### #2 heat sensitive
The next most recessive allele codes for a heat inactivated version of tyrosine
meaning that the final colour of the rabbits fur will depend on the temperature
conditions it was raised in, and also that within a certain temperature range
there will be a distinct colour difference between the warmer and colder parts
of the animal.

###### #3 localised functionality
The next most recessive allele codes for a version of tyrosinase does not fold
quite "correctly" and as such does not glycosalate in the same way as the
dominant version. melanocytes which produce the grey pigment will still accept
this tyrosinase into their cytoplasm and as such grey pigment will be produced.
however as intracellular transport is affected the tyrosinase either cannot
reach or cannot enter melanocytes capable of producing the brown pigments and so
they rabbits are full grey

###### #4 fully functional
the most dominant allele codes for a fully functional form of tyrosinase and
the combination of grey and orange pigments produced gives the rabbits a full
brown coat.

#### human Blood types.

#### details
human blood types are determined by one gene locus with three alleles.
allele O is recessive, and codes for no functioning protein.
allele A and B are co dominant and each code for different sugars which are
attached to the outside of red blood cells.






#### Bombay phenotype

##### details
parents A and AB, child O, mates with A, results B, AB and O.

##### explanation
A and B sugars are attached to red blood cells by appending them onto the end
of glycoproteins called Substance H which are already attached to the cell.
If a individual is homozygous recessive for substance H genes, none will be
produced and so even if the have A or B alleles they will have a functionally
O phenotype.

#### sex chromosomes
 lethal alleles, or alleles linked with low fertility located on the Y
chromosome in humans are very rare.
disease related to homozygous XX are more likely to be suffered by woman, as
they will have one working copy of the given gene so they can survive but
one defective copy which leads to the illness. males on the other hand will
either survive and be entirely healthy with a working copy or die, if they are
even born if they have only a defective copy.




##### lethal dominants
lethal dominants may occur in one of two cases. First, if the onset of the
disease/ defect related to the Allele only occurs after reproductive age.
Secondly if a De Novo (random, once off) DNA mutation occurs in a given
individual.
lethal dominants can also result from an Allele which interferes with the
formation of aggregates of the protein which the gene codes for, or receptors of
that protein,or finally if it has a detrimental effect on another protein or
gene (such as the yellow mice)

#### examples

##### Yellow Mice.

##### details
Y allele codes for the yellow coloured pigment.
y codes for other pigment/no pigment.
Y is a lethal Dominant(?) Allele. As a YY individual with not develop bast
the early stages of implantation
into the uterine wall.

##### molecular/genetic explanation
When the Y allele is present at the gene loci it causes a deletion in which,
its own gene promoter, as well as the coding sequence of the gene upstream
of it(the MERC gene necessary in RNA processing) are removed. the net result
 being that proteins relating to the gene upstream are no longer removed, and
 the pigment gene is now promoter by the MERC promoter instead of its own
promoter

##### Result and Phenotype
the result is that Yy individual still produce enough RNA processing proteins to
survive/ avoid miscarriage,but they are more susceptible to diseases such as obesity
and cancer. Also their yellow pigment gene is promoted heavily during early
development leading to yellow coloured fur.
YY miscarriages early in pregnancy as it cant produce necessary RNA processing
proteins.
yy is health individual without yellow colour.
birth rations will be skewed by the fact that YY individuals are not actually
born.
NOTE: this is an example of incomplete dominance of Y with respect to MERC.

### Effective Of the environment on genetic expression.
#### examples

Myopia(short sightedness):

Populations where book learning is not inforced/encouraged from an early age do not show high rates of myopia even though the genetic dispostion between the two populations is very similar.



### Phenylketonuria
single mutation
can affect brain development however if the correct diet is followed then all detrimental health consequences can be ignored.

### Diabetes type II
lifestyle disease, the genetic component is estimated to change the likelihood of developing the disease by 2%, but the main causative factors of the disease relate to diet.

### Lactose intolerance.
Individuals with lactose intolerance cant digest lactose, however the bacteria in their gut van which leads to the build up of gas. this can be very painful for the person in question.In humans lactose intolerance would naturally occur after weening however as cow milk is a staple part of most human diabetes the genes controlling lactase production are never switched off and the ability to digest lactose is not normally lost.

### Himalayan Phenotype.
### Cancer

#### basics
A cancer is a group of cells whose proliferation is uncontrolled, and which can spread to other locations in the body which are normally populated bu other cell types.

##### benigin tumours
grow but do not spread.

##### malignant tumours
over proliferate and invade other body tissues/ areas in the body.  

#### types of cancers

##### Carcinomas
 derived form epithelial cells

##### Sarcomas
 derived from connective or muscle tissues. Known as Osteosarcoma in the case of bone tissue and Kaposi sarcoma in the case of soft connective tissue.  

##### Lymphoma and leukaemias
1. Cancers of the hematopoietic system.
2. Lymphoma (solid tumour)
3. Leukaemia excess of circulating immature blood cell precursors.
##### Cancer of the nervous system
brain and central nervous system.

NOTE: most common in children

##### Germ-line cancer.

##### Cancers of breast, prostate, lung, pancreas, and colon.

### Causes of cancer.
In most cases cancer is not inherited. In  breast cancer/ovarian cancer BS1, BS2 play some role in genetic disposition, perhaps 5-10%

### Hereditary.
Cancer normally sets in during old age as cells must accumulate a series of a specific set of notations, (exactly which set of mutations accumulated lead to cancer is highly individual)
#### Environmental factors
1. UV
2. X rays
3. Alcohol
4. Overcooked food (hetercyclic amines,polycyclic aromatic hydrocarbons-meat. acrylamide- potatoes
5. Azo dyes
6. tartrazine (food colourants)
7. Nitrate cured foods
8. Pesticides

#### Cancer initiators
agents which cause DNA damage (mutagens). these factors may be chemical, biological (such as HPV- human papilloma virus cause latent genetic damage- Cancer predisposition.

#### Cancer Promoters

Promotes excessive proliferation (does not directly damage DNA).

##### Examples
1. wounding
2. phorbol esters
3. HRT/ oestrogen (breast cancer)
4. hepatitis B (promotes stomach cancer)
5. HIV (Kaposi sarcoma, this disease is always present but only manifests itself in immunocompromised individuals)

#### stages of cancer development  
1. normal cells (mutation occcurs)
2. increased ability to proliferate (mutation)
3. 1-2mm tumour of rapidly proliferating cells which do not undergo apoptosis of differentiation. (mutation)
4. vascularized growing tumour (mutation)
5. large tumour capable of invading near by tissue (metastasis)

NOTE: cancer results from a series of somatic mutations, affecting the same cell, and different cancers are genetically heterogeneous.

#### Genetic instability in cancers.
deficient local DNA repair leads to the accumulation of point mutations increased chromosomal instability and gross genome abnormalities.

#### summary of key properties
1. disregards ex/in growth regulation signals.
2. avoid apoptosis, differentiation and replicative senescence
3. genetically unstable.
4. invasive
5. metastatic (survive and proliferate in foreign) sites.

#### Penetrance
 percentage of the population who demonstrate at least some degree fo phenotypic expression.

#### Expressivity
reflects the range of expressions of the gene/allele present in the population.  

#### Genetic complementation
two different mutations in heterozygous condition affecting the same protein/ pathway can compliment each other to cause a novel phenotypic effect.

##### examples
rare form of albinism: both mother and father were normally pigmented and there families had minimal instances of albinism. however both children where albino' s . Analysis found that this albinism stemmed not from those genes and mutation most commonly associated with albinism but rather from a combination of recessives (heterozygous) from both father and mother (mutations in TYRP1 gene), which on their own would have little or no effect on the pigment production pathway, but when combined reduced pigment production significantly.

NOTE: Rufus albinism leads to a phenotype with reddish hair, lighter skin, and blue grey eyes.

#### Forked line probability method.
 
##Chromosome level
(revise mitosis and meosis)
copy slides
chromosome basic structure
#### chromosome groups

##### meta-centric
centromere is half way up the length of the chromosome

##### submetacentric
centromere is more to the one side of the chromosome than the other,

##### Acrocentric
the centromere is very far to the one side of the chromosome, with a long arm containing most of the genes and a short arm containing predominantly temolmeric DNA.

##### Telocentric  
humans do not posses any telocentric chromosomes, but certain insect of crustation species do.

##### Holocentric
Centromere like structure exist along the entire length of the chromosome. this may decrease the chances of faulty division/ segregation.

##### chromosome banding
bands were named and used to locate specific genes. the banding patterns are due to uneven DNA densities in the coiled structure of the chromosome.

##### Size and shape of different chromosomes.

### Chromosome level mutations

#### Aneuploidy (Spelling)
Each cell is has at least one extra chromosome or is missing at least one chromosome

##### Trisomy
when an individual inherits three copies of a particular chromosome.
In Humans only three of all the possible trisomies are viable, (as in individuals with these mutations will still be born and not terminated during pregnancy)

of these three (Trisomy 13, 18, and 21) only individuals with trisomy 21 can survive past the first few years of childhood. Individuals with Trisomy 21 have down syndrome.

###### Down syndrome (Trisomy 21)
Frequency: 1/1000
Effects
1. mental retardation
2. short stature
3. heart disease
4. shortened life span.

###### Edward Syndrome (Trisomy 18)
Effects:
1. kidney and intestinal malformation
3. heart defects.
4. mental retardation

NOTE: only 8% survival past the first year.

###### Patua syndrome (Trisomy 13 )
Effects:
1. Kidney and hear defects
2. polidactily (too many digits)
3. Nervous system abnormalities
4. Death within the 1st year.



NOTE: the smaller relative size of chromosome 21, meaning that it contains less important genes may be related to its increased viability .

##### Monosomy
when an individual inherits only one copy of a particular chromosome.

##### Nullisomy
having no copies of a particular chromosome.

#### Polyploidy

#### Structural rearrangements
1. deletion
2. duplication
3. translocation
4. inversion
5. fission
6. fusion

#### non disjuncture
(review mitosis and meiosis)
If non disjuncture happens in meiosis it is generally worse than if it happens in mitosis as it will affect the whole organism.

##### Down Syndrome
occurs in about 1 of a thousand live births. originates from non disjunction in the egg cell rather than the sperm cell because the arrested development of the egg cells leads to decay of the separating and marking proteins as well as the spindle fibre leaving the whole process open to more error.
8% of individuals survive for one year ?

##### turner Syndrome monosomy X
Frequency 1/20 000 (only occurs in females )
Effects:
1. webbed neck
2. short stature
3. underdeveloped ovaries and lack of secondary sex characteristics.

##### Klinefelter syndrome XXY
Too many X's ,XXY , XXXY, XXXXY, XXXXXY. (XXYY)
Effects:
1. Decrease in testosterone levels
2. the more X's the higher the chance of brain damage.

##### XYY
XYY has the phenotype of a normal make.
men with 2 Y chromosomes tend to be tall as there is a cumulative effect adding to height. (they may also exhibit increased aggression but this in unclear)

##### XXX
Frequency: 1/1000
Effects:
1. Normal female with normal fertility.

NOTE: this conditions is seldom diagnosed.


#### Polyploidy
Three or more complete sets of chromosomes present in somatic cells.
#### 2n-20n
caused by non reduced gamete fomration, which is very rare in males and non viable.

most often flowering plants are polyploid, they are even specifically breed to have more chromosome sets as this usually increases fruit and flower size.


##### Examples

water melon
must be bread from a tetraploid(?) and a diploid to get a tripliod
infertile plant.

Kiwi fruit
12-16 copies

Strawberries
4 copies (tetraploid)

Frogs
_Xenopis levis_  4n
_Xenopus tropicalis_ 2n

NOTE: plants are better adapted to polyploid because:
1. they are not as confined to a set physical form so different in growth and development genes operation levels are not so important
2. they can reproduce vegetatively so polyploid individuals aren't as severely evolutionarily disadvantaged.
3. less precise sperm targeting is necessary.



#### Deletions
deletions often occur by chromosome breakage during the cell cycle. These deletions may be terminal of interstitial. If the centromere is lost, the
entire chromosome will be lost.

##### Piece of Chrome
sometimes multiple genes are lost when both strands break and a part of the chromosome is permanently lost. A specific deletion on chromosome 5. (where the entire p arm is deleted leads to a serious syndrome)

##### Cri du chat (5p-)
Frequency: 1/20000-50000
Effects:
1. Intellectual disability,
2. low muscle tone
3. microcephaly
4. distinctive facial features.
5. normal life expectancy  

#### Duplications
Cause: retro transposition, or non-allelic homologous recombination in repeat rich areas.

A small region of a chromosome is repeated, and placed next to the existing copy (in tandem). duplications tend to happen in high repetitive sequences of DNA. These repeated regions mean that the chromosome will note line up very well with its homologous pair.

passed on by unequal order, and may or may not have a phenotypic effect.


#### Huntington Purea
Repeats alter gene function, which can lead to brain damage.
Repeats may have a positive effect because they allow one copy of a gene to evolve independently, specialise in a different function while the original function is still conserved by the other copy of the gene.

#### Inversions.
Cause: non-allelic homologous recombination
Effects:
1.  pairing in meiosis occurs via looping
2. cross over in the looped region can cause major deletions or duplications.

#### Translocation
Effects:
1. novel chromosomes are generated
2. individuals are usually infertile.
3. can result in aneuploidy (such as 14q21q inherited down syndrome)
4.

##### terminal
A piece breaks off a chromosome and sticks to the end of a non-homologous chromosome.

##### reciprocal
Exchange of pieces between non-homologous chromosomes
#### Robertsonian

two Acrocentric chromosomes lose their short arms and get stuck together.

#### Fission
deletion (loss of acentric piece)'

#### Fusion
bicentric chromosomes (breakage in mitosis)

##### breakage fusion bridge cycles in cancer.
Copy image.
1. end of a chromosome breaks off (during meiosis)
2. one daughter cell inherits the chromosome lacking a telomere.
3. the new cell enters S phase and replicates its DNA (it now has two sticky DNA ends right next to each other.
)
4. loose DNA strands bind to form continuous DNA loop.
5. fused sister chromatids are pulled apart during mitosis, causing breakage at a new sight
6. one daughter cell inherits chromosome with duplicated genes and again a free sticky end. this cycle can then begin again leading to even further chromosome distortion.

#### Imprinting

##### Hinny vs mule

###### Mule
horse mother donkey father
hardy and obedient

###### Hinny
donkey mother, horse father.
temperamental, untrainable.

##### Intersistial deletion of chromosome 15

###### Prader -Willi syndrome
Cause: deletion from father, ZNF127 and IPW inactivated on the maternal copy.  
effects:
1. mental retardation
2. obesity
3. diminished growth.

###### Angelman syndrome
1. deletion from mother: UBE3A inactivated on paternal copy.  
2. mental retardation
2. epilepsy
3. lack of motor development.

#### X inactivation
the inactivation of one of two X chromosomes in females via DNA methylation.

NOTE: which X chromosome is inactivated varies randomly from cell to cell.this inactivation process turns the inactive X into a bar body, (highly methylated and rolled up in histone proteins).

#Unsorted notes

1. Himalayan Phenotype.
2. Size and shape of different chromosomes.

Dr Fanucchi, GH416, 
Date: 17/06/2018.

# Course overview. 
ezymes and function \newline 
Garret and Grisham Biochemistry.\newline 
Questions in class (write it down)\newline 
*Prelab test before every practical* \newline 
(lecture + prac but only theory relevant to prac, no writing up prac reports. ) \newline 
lab etiquette (dress+ approporitate work+ sharing with lab partner, fall behind group, spot questions in the lab, prac ends at 5 sharp)\newline 
*Seating plan new, on Sakai*\newline 
Changes from prach manual (check course outline, Friday WSS4, 423-426 text book)\newline 
tues+wed=P115, thurs=WSS4.\newline  
*No PH prac*.\newline  
Due date of essay: *8th August*. (submitted electronically+ marked out)\newline 
*Submit essay in box 1:30 10th August, must print in turn it in format, 15\% or less*\newline 
Poster is due on a Monday, submit to Sakai, \newline
31 Aug, Class test 2-4pm OLS +GH403. \newline 

## Dr Fanucchi class rules
 #. Own studying. 

## Assignment. 
The MIND diet, med+DASH (DASH for hyper tension+neuro disease), fish altsheimers, weight loss, 2 veg per day... 
Argumentative essay, critically evalute (find actual 2015 paper, met pathways+ metabolism+ for or against but must take a stand)

### Essay format 
Three sections+ tittle (intro, body, conclusion), 

#### Intro 
must but work into context, it must give background info however it must be very focussed on the concept at hand. Must make clear the relavance and significance of the subject. Can finish with a sentance/2 disscribing structure of the essay ("In this essay i will") , very end must have main claim (must be _interesting_ and specific, must make absolutely clear for or against + angle),

##### Main claim
Short, consise, specific, clear and easy to recognise (ie at the end of the intro) \newline 

#### Body 
Set of paragraphs, moving from more general to specific, broadest info first to more specific. Includes reasons in support of claims. Give reasons + cite evedence from the literature.\newline  
(Could attempt to include a rebuttle, couterarguments)\newline 

##### Transition sentances 
link paragraphs to imporve flow. 

##### Topic 
topic sentance for each paragraph. 

#### Wrap up sentance
Relates the topic back to the main claim. 

NOTE: Lots of citations, remember writing fellows. 

### Final notes
 #. 800-1100 words, 
 #. own research + interpetation of literature,
 #. must be submitted through turn it in. 
 #. print correct format.
 #. print + sttatch marking rubric 


#### Conclusion 
Restate the main claim and its importance. Summarize each of the main points in the claim again. 

*NB*\newline 
No new information, and no references. 

# Scientific Poster 
Commonly presented at scientific posters, (ppl present current research, orally and in posters, advantage is that it is very consice)\newline 

NOTE: \newline 

 #. Not too much text. (NB)
 #. read 1m away (A1 size) 
 #. lots of visuals. diagrams etc. 
 #. The take home message must be clear (stance must be clear)
 
NOTE: 
 #. 4pm monday, 
 #. pdf format
 #. 594x 841mm. \newline 
 
# Background 
Governed by relatively small set of relatively simple rules. 

# Chapter 13 Biological catalyst. 
enzmyes in metablosim, accelerate and control biochemcial reaction rates. In specific speed up dergradation of nutrients adn generation adn tranfromation of metabolites. Only very specialised regulatory enzymes control biochemical reactions, often found at juncture points. Suhc enzymes change the rate depending on the cell's metabolic requirements. (ie more require speed up sythesis and slow degradation and visa versa.)

Many biological reactions produce a lot of energy but too slow, enzymes exert kinetic control overthermodynamic potentiallity. (kenetic as in rate and speed)\newline 

## Distinctive features

### Catalytic power. 
how good it is at being a catalyst/ speeding things up 
Calculated as the ration of catalysed to uncatalysed reaction

### Specificty. 
Acts on a specific substrate, very selective no promiscuous. ()
selectarivity for substrate. Cannot change the eqillibrium position, enzymes are very selective. One advantage is less/no biproducts. 

Achieved by secific shape adn charge, (complimentarity)

#### Yield 
Yield decreases at each step in a multi step process, (yield is decreased at each step by the formation of byproduct)

### Regulation 
Not all enzymes, key enzymes, maintain rates at appropriate levels for the cells requirements. 

### Mild conditions 
happed under very mild conditions (pH and temp), as proteins suseptable to denaturation, furthermore often 


```{r, echo=FALSE}
# example 

(3*10^4)/(3*10^(-10))

```
## Background 

### Nomenclature
Enzymes are grouped into 6 different categorised based on reactions which they catalyse. 

#### Enzyme commision number. 
Furthermore each enzyme is given an enzyem commision number, this is a standard however a "common name" is also often used. 

#### Oxidoreductases. 

#### Transferases  

#### Hydrolases.
Cleaves in presence of water. 
(Most common type)

#### Ligases. 
form bonds 

#### Isomerases. 


#### Lyases. 
break bonds.
 
NOTE: Check in notes if all enzyme groups area present. 

```{r,echo=FALSE}
#Excersize

```

### Principles of chemical kinetics. 
$A\rightarrow B$ \newline 
assume spontaneous and irreversible. 

rate defined in terms of change in concentration of the reactants or products. 

NOTE: Copy notes from chemical kinetics. 

$v=-\frac{d[A]}{dt}=\frac{d[P]}{dt}$ the infinitesimal rate. 

### Rate Law. 
$v=\text{k}[A]$ where k is the rate constant. 
reverse rate $k^{-1}[P]$\newline 

$V\propto [A]$ \newline 
k is the rate constant of the constant of proportionality.



# Enzymology 

## Terminology 

### Reaction order 
The reaction order with respect to a specific component is the power to which that component is raised in the rate equation for the reaction. 

The overall reaction order is the sum of the reaction order's of all reactants. 

NOTE: the reaction order is not expressed as an integer (such as 1 or 2) but rather as first, second...etc. 

### Molecularity 
The molecularity of an (elemental) reaction is the number of molecules which must interact (symultaneiously) during the reaction. 

NOTE: the molecularily is a (positive) integer number(1,2...) but the reaction is then discribed as: unimolecular, bimolecular,etc. 

## Progession curves 
PRogression curves show the change in conentration of a reactant/product agaisnt time. 

## Initial rate 
The initial rate $v_0$ in the reaction rate at t_0. 

NOTE: an increase in intital reactant concnetration leads to a higher inital rate. 

## Rate law 
The rate law for a reaction: \newline 

$a_1A_1+a_2A_2...\rightarrow b_1B_1+b_2B_2...$ \newline 

is given as: \newline 

$v=k[A_1]^{n_1}[A_2]^{n_2}$\newline 

where $n_1,n_2...$ are experimentally determined constants. 

NOTE: In the case of an elemental reaction, the exponents will be equivalent to the equilibrium coefficients. 

## Rate 
For the (elemental) reaction: \newline 

$a_1A_1+a_2A_2...\rightarrow b_1B_1+b_2B_2...$ \newline 

The instantanious rate of dissapearance of $A_1$ is given by:\newline 
$\frac{d[A]}{a_1dt}$ 

This rate can be related to the rate or dissapearance of other reactants and products as follows:\newline 

$v=-\frac{d[A_1]}{a_2dt}=-\frac{d[A_2]}{a_2dt}=\frac{d[B_1]}{b_1dt}=\frac{d[B_2]}{b_2dt}$ 




i## Enzyme Kinetics. 

### Maximum velocity 
Maximum catalystic rate\newline 
Donated by $V_{\text{max}}$

### Binding affinity
the strength of the bond/ enthalpy of bond formation \newline 
dontated by $K_m$

### Mechalis mention. 


#### Derrivation . 


##### Assumptions 
 #. The back reaction (E+P to ES in negligable.) $k_{-1}>>>k_2$ (Mechalis menten) 
 #. THe concnetration of the ES remains constant at a steady state (steady state assumption.) (Briggs- Huldane. )
 
 The inital velcoity can be expressed as a function of substrate concentration. 
 
 Total enzyme concentration: $[E]_T=[E]+[ES]$\newline ($[E_T]$) si known from the amount of enzymes added. 
 
$v=k_1[E][S]$ \newline 

$-v=k_{-1}[ES]$ \newline 

$k_2[ES]$ \newline 

By the constant rate approximation. 

$k_1[A][S]-k_{-1}[ES]-k_2[ES]=0$ \newline

(Look up coniditions or steay state assumption. )


$\rightarrow k_1[E][S]=k_{-1}[ES]+k_2[ES]$ \newline 
as $[E]_T=[E]+[ES]$ newilne
$\rightarrow k_1([E_T]-[ES])[S]=k_{-1}[ES]+k_2[ES]$
$\rightarrow \frac{([E_T]-[ES])[S]}{[ES]}=\frac{k_{-1}+k_2}{k_1}=k_M$ 

As $k_2>>>k_{-1}$\newline 

$(\frac{k_{-1}}{k_1})$ 

NOTE the unit of $K_m$ will be $mol\cdot dm^{-3}$


$\frac{([E_T]-[ES])[S]}{[ES]}=k_M$


$\frac{[E_T][S]}{[ES]}-\frac{[ES][S]}{[ES]}=K_M$ (solve for [ES])

$[ES]=\frac{[E_T[S]]}{K_M+[S]}$

Alsso as 

$v=k_2[ES]$ (rate o prouct formation). \newline 
$[ES]=\frac{v}{k_2}$ \newline 

$\rightarrow \frac{v}{k_2} = \frac{[E_T][S]}{K_M+[S]}$\newline

$\rightarrow v= \frac{k_2[E_T][S]}{K_M+[S]}$ ($[ES]\approx[E_T]$ for excess substrate) \newline 

$\rightarrow v=\frac{k_2[ES][S]}{K_M+[S]}$ $v=k_2[E_T]=v_{max}$ As ($[E_T]=[ES]$) for saturated solution. 

$v=\frac{v_{max}[S]}{{K_M+[S]}$ where v is the initial rate. 

$K_M$ and $v_{max}$ ca be used as general measures of catalytic ability. 

$K_M=[S]$ at \frac{v_max}{2}

$v=\frac{v_{max}[K_M]}{{K_M+[K_M]}=\frac{1}{2}v_{max}$ 

product concentration against time can be plotted and the tangent taken to give v at a given t. 


NOTE:

 #. either ony one susbstrte or concnetrtio o other substrates held constant. 
 #. E is in steady state. 
 #. reaction is irreversible., or 
 
 Copy down assumptions!!. 
 
 
 $\text{ln}(v)=\text{ln}(\frac{v_{max}[S]}{{K_M+[S]}})$
 
 $\rightarrow =$
 
NOTE: $v_{max}$ can be approximated}

## Thermodymanics. 
Thermodynamics is the study of the flow of energy.
Energy is obtained from food and utilised to perform biological functions.
bioenergetics of life. 

### Spontenaity. 
Thermodynamics can be used to determine which metaloic reactions will happen spontaneously. 

### Background 

#### System 
The portion of the universe which is of particular interest. 

##### Isolated 
The system can not exchange matter or energy with its surroundings. 

##### Closed 
Energy but not matter is exchanged with the surroundings. 

$$$$$ OPen 
Both energy and matter canb be exchanged with the surroundings. 

$$ Thermodynamics laws. 

### Fisrt Law.
The total energy of an isolated system is constant. 

### Second law. 
The total entropy of an isolated system cannot decrease but can only remain constant in the case of a system in eqillibrium or undergoing reversible reactions, or increasing when the system is undergoing reversible processes. 

## Internal energy 
The sum of all molecular kinetic (virbtarional rotational and translational) and potential mocecular energies. INternal energy is a state function so is independant of the path taken to reach it. 

$$$ Change
The internal energy of a system may change with the transfer of energy as heat or work. 

$\Delta E=w+q$ 

NOTE: The units for heat work and internal energy are joules. 

#### Aquisitive convention. 
w is positive when doen on the system. heat is positive when heat is transfered to the system. 

## Work 
PRessure volume work. $w=-p_\DeltaV$

$$$$Surroundings
The surroundings are everything other than the system. 

i## Enthalpy 
At constant volume $\Delta E=q_v$

$H=E+PV$ by definition. 

#### Change. 
$\DeltaH=\DeltaE+p\Delta V= q+w+P\Delta V= q-P\Delta V+P \Deta V =q$ (As pressure is constant in a biological environment)
$H=q_p$

Volume changes are minimal in biological systems (as reactions are not often in gasseous form) so $\Delta E \approx \Delta H$

$\Delta H$ is measured suing a calorimeter. 

$\Delta H ^\circ = -R \frac{d(lnK_{eq})}{d\frac{1}{T}}$

THe Van't Hoff plot gives delta H. 

On the y $Rln{K_{eq}}$ is plotted against $\frac{1}{T}$ on the x axis. so that the sslope =$-\Delta H$
Endothermic is the absorption of heat. 

NOTE: the denaturation of an enzyme is endothermic becuase of H-bond breaking, and exposure of hydrophobic groups. 

## Enzyme kinetics. 

### $K_m$
For an enzyme catalysted reaction there is first order kinetics for low concnetrations of substrate, tending towards zero order kenetics as substrate is increases to the point of enayme saturation. 
$K_m$=\frac{V_{max}}{2}$

A high $K_m$ indicates tht a lot more substrate is necessary for the enzyme to become saturated, hence it indicates a low affinity for the substrate. 
A low $K_m$ indicates that the enzyme bcomes saturated easily and has a high affinity for its substrate. 

#### favoured substrate. 
The enzyme with the lower $K_m$ will be favoured. 

$$$ Turnover number/kinetic effeciency
Reffered to ass molecular activity ($K_{cat}$)  the turnover number indicates the maximal catalytic acitvity, defined as the number of substrate molecules converted into product per enzyme per unit time under saturating substrate concentrations. 

for the reaction $E+S \rightarrow ES \rightarrow E+P$  at initial conditios $K_{-2}$ is zero and $K_{cat}=k_2$ 

At saturationg substrate concentrations $v=v_{max} = K_2[ES]=$K_2[E_T]=\frac{v_{max}}{[E_T]}=k_{cat}$
NOTE: $k_{cat}$ is a first order rate constant (with units $s^{-1}$)

$k_{cat}$ is only informative at very high substrate concentrations which ins't physiological. (In the body normally $\frac{[S]}{K_M}<1$)
Vut we knwon $V_max=k_{cat}[E_T]$
So from mecahlis meten $v=\frac{v_{max}[S]}K_M{+[S]}$

As $[S]<<K_m$ 

$v=\frac{v_{max}}{K_M}$

At very low substrate concentrations $E_T$=[E]
So $v=\frac{k_{cat}}{K_m}$[E][S]

$\frac{K_{cat}}{K_m}$ is second measure of effeciency. It is the apparent second order rate constant for the reaction of enzyme and substrate to give poroduct.  
The higher this measure the better the kinetic effeciency and the berrer the affinity at low {$S}$ , i.e. the higher the catalyti effeciency of the enzyme. 

#### Catalytic perfection 
Enzymes which reach catalytic perfection are limited only by how fast the enproduct and enzyme can diffuse together. Rate of diffusion of enzyme and substrate in solution.  (maximum value is about $10^{8-9}$)

NOTE: $\frac{K_{cat}{}K_m$ applies to the initial region of the [S] vs rate graph, whereas $K_{kat} reprisents the later region onces $[S]^0$ . $v_0$=k_{cat}[E]^1[S]^0
$V_0=\frac{k_{cat}{K_m}}[S][E]$

## Enzyme inhibition 

### reversible 


### reversible
binds non eqilibrium 

#### Compeditive 
Inhibitor binds only to the free enzyme (so has to bind (often )) to the actvie site and so must look like the enzyme. 
Compeditative inhibitors increase $K_m$ by lowering the effective substrate concentration. 
$V_{max remains constant}$


#### uncompeditative 
THe inhibitor will only bind to the enzyme substrate complex. $V_max$ drops as EI complex formed which lowers [E] $K_m$ also decreases (because of the law of mass action, leads to an increase in affinity. 

Line weaver burk is a series of parallel lines. $\frac{V_max}{}K_m$) the slope, does not change. 

#### non compediitve. 
Can bind to enzyme enzyme substrate or both 

##### Pure 
When the affinity for the enzyme or the enzyme substrate complex is identical it is ppure

##### Mixed 
IF the affinity for the enzyme and the enzyme substrate complex differs then the enzyme is mixed. 

$$ Bisubstrate reaction 
Two substrates interact with the enzyme 

$$$ Single displacement/ sequential 

### Random 
There are two substrate binding sites, so substrates can bind in any order. 

### Ordered
Binding of one substrate inductes the formation of a second substrate binding site by a conformational change within the enzyme its.


### Double displacement.
Follow the pingpong method. The first substrate binds and covalently modifies the enzyme (leaves part of itself behind) . The second substrate can then bind, takes the part left by the first substrate onto itself and is then released. 

## Thermodynamics second law. 
The entropy of an isolated system can only decrease or remain constant (in eqillibrium or when reversible processess aer taking place. )Entropy reprisents energy dispersion. 

$S=kln(W)$
where w is the number of microstates

Moving from disorder to order requires energy input in the form of work. (this is of particular importance in a biological system where reactions, at least catabolic reactions involve the formation of highly ordered macromolecules. 

### Third law 
S=\idsplaystyle\int ^T _0 C_pdlnT )$ the absolute entropy of a perfect cystal at zero degrees kalvin is zero. . 


### Heat capacity 
$C_p=\frac{dH}{dT}$

the heat capacity is the amount of heat required to increase 1 mol of substance by one degree. 

#### $\Delta C_p$ 
Positive change implies more heat stability. which implies that the molecules have new ways to move so more enrgy can be added for a temperature increase. Conversely if the change in $C_p$ is negative there are less ways of motion. 

### Free energy 
The free energy gives the spontenaity of the reaction

$\Delta G=\\Delta H -T\Delta S$ when change in gibbs free energy is zero the enthalpic and entropic compoents are equal. at this point the reaction is at eqilibrium. A spontaneous/ exogonic reaction has a negative $\Delta G$ 

NOTE:remeber the $\Delta G$ indicates the enrgy change between the reactants and products. 

If $\Delta H$ is positive the reaction is said to be enthalpically opposed, or in the case of  negative $\Delta S$  the reaction is entropically opposed. (the opposite of opposed is favoured. 

Free energy entropy are temperature dependant, so the temperature range of the experiment must be specified. backspace32

#### for non standard conditionbackspace32

$\Delta G= \Delta G ^\circ RTln(\frac[C][D]){[A][B]}$

$\Delta G =-RTln (K_{eq})$

In biological conditions we use a modified standard state $\Delta G^\circ '$ which means pH is at seven .


NOTE: a comparison of several thermodynamic parameters can give a reasonable idea of the reaction, one on its own is not particularly useful. 

ie for protien unfolding, $\Delta G$ is positive, $\Delta S$ will be positive. 


## Analysis of processes according to thermodynamics. 
Determine entropically and enthalpically opposed /favoured. If a reaction still occurs it can be called enthalpically/entropically driven. 
Th dissolution of a polar substance in water leads to a decrease in heat capacity whereas the dissolution of a non polar substance increases the heat capacity, this is becuase during solvation the water molecules form a solvation/hydration shell around polar substances reducing the freedom of their moviment. 
In the case of hydrophobic molecules water formes an ordered cage, clathrate cage, around the hydrophobic molecules, interacting with eachother as a pose to the hydrophobic molecule. The clathrate is more laible (that is more free to exchange hydrogen bonds. Liable waters result in a positive change in heat capacity because they have more ways to store energy. )

Protein folding. There are more hydrophobic groups exposed as a protein unfolds, leading to increase in the $C_p$, conversely when a portein folds these groups are removed from solution and therefore the heat capacity decreases. (also freedom of movement of the protein) 

NOTE: the hydration shell is not just the water dirrectly in contact with the ion but also in layers outwards as if the shell orientates one way then the outside ones also orientate with respect to those. 

NOTE: remember all of the thermodyamic parameters are constant. 


NOTE: Pracs start at 2, test covers everything done in enzymes connetcs, but focussed on the lab and there is lab prep work. 

NOTE: water will formed orded shells areound charged ions making them more ordered. 



NOTE: remember $\Delta G$ gives an indication of the eqillirium position. 


### Biological energy flow. 
The vast majority of energy derrives from the sun 

#### phototrophic organisms. 
Absorb energy from the light energy directly and store it in the form of various organic molecules. 
Energy is released in cotrolled exergonic reactions taht drive other reactions requiring energy (endergonic reactions)(i.e. there is an energy flow between exergonic and endergonic reactions. Mediated by a small number of biomolecules. (high energy). chemical energy is released in controlled spontanious reactions driven by  exergonic reactions that drive life. High energy molecules react spontaneously with a substance in the environment with high $\Delta G ^\circ '$ values,releasing large amounts of energy. The environmental molecules are most mcommonly oxygen, the most common high energy organics are the reduced coenzymes. $NADH, NADPH, FADH_2$. Water is another common environmental molecule, which phosphate compounds (ATP) of thioesters, such as acetyl coA. 

NOTE: ATP does not have the highest energy of hydrolysis or the lowest, with regard to the energy shuttle mollecules, ie it is intermediate. This is important as it can take nergy from very high energy phosphates and release energy to from lower energy phosphate compounds. Because energy can be tranferred by ATP in this way it is known as the energy currency of the cell. 


### Chemoheterotrophes. 
Comsume the organicm molecules and release the stored energy in a series of oxidative reactions. 

