\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother


\date{}

\begin{document}

\hypertarget{genetics}{%
\section{Genetics}\label{genetics}}

\hypertarget{course-outline}{%
\subsubsection{Course outline}\label{course-outline}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  extension of Mendelian genetics
\item
  Chromosomes structure and changes
\item
  Vertebrate development
\end{enumerate}

\hypertarget{interactions-between-genes-alleles-and-phenotypes}{%
\subsection{Interactions between Genes, Alleles, and
Phenotypes}\label{interactions-between-genes-alleles-and-phenotypes}}

\hypertarget{incomplete-dominance}{%
\subsubsection{Incomplete dominance}\label{incomplete-dominance}}

\hypertarget{genetic-molecular-cause}{%
\paragraph{Genetic/ molecular cause}\label{genetic-molecular-cause}}

The recessive allele does not produce any functioning protein. The
dominant allele does function protein but not in high enough quantities
to result in the same phenotypic effect as a heterozygous dominant

\hypertarget{phenotypic-effect}{%
\paragraph{phenotypic effect}\label{phenotypic-effect}}

A heterozygous individual will display a phenotype intermediate to that
of homozygous dominant and homozygous recessive

\hypertarget{dominance}{%
\subsubsection{Dominance}\label{dominance}}

\hypertarget{geneticmolecular-cause}{%
\paragraph{Genetic/molecular cause}\label{geneticmolecular-cause}}

The recessive allele does not produce any functioning protein. The
dominant allele does function protein, which it produces in comparative
amount to a homozygous dominant individual

\hypertarget{phenotypic-consequence.}{%
\paragraph{Phenotypic consequence.}\label{phenotypic-consequence.}}

the heterozygous phenotype is identical to that of the heterozygous
dominant individual.

\hypertarget{co-dominance}{%
\subsubsection{Co-dominance}\label{co-dominance}}

\hypertarget{genetic-molecular-cause-1}{%
\paragraph{Genetic/ molecular cause}\label{genetic-molecular-cause-1}}

Both alleles are transcribed/translated to form a functioning protein

\hypertarget{phenotypic-effect-1}{%
\paragraph{phenotypic effect}\label{phenotypic-effect-1}}

The effect comes from the full expression of both alleles

\hypertarget{dominance-hierarchies.}{%
\subsubsection{Dominance hierarchies.}\label{dominance-hierarchies.}}

It is often the case that there will be a number of alleles for one
gene. In which case a dominance hierarchy is usually present such that
allele A is Dominant to B is dominant to C and so forth (ie dominance in
a hierarchy is transitive)

\hypertarget{pleiotropy}{%
\subsubsection{Pleiotropy}\label{pleiotropy}}

Genetic/molecular cause One gene locus affects the expression of several
characteristics. Phenotypic cause Phenotype of one characteristic will
change in conjunction with the phenotype of another characteristic

\hypertarget{notes-about-colour}{%
\subsubsection{Notes about colour}\label{notes-about-colour}}

As many genetic examples are based on colour it is worth understanding
the basics of colouring in animals. colour results from pigments which
are proteins, individual genes can either code directly for the
production of these proteins or as an intermediate/enzyme which aids in
the production of these proteins.

\hypertarget{white}{%
\paragraph{White}\label{white}}

white does not result from a specific pigment but rather from a lack of
pigment.

\hypertarget{co-dominance-and-incomplete-dominance}{%
\paragraph{Co-dominance and Incomplete
dominance}\label{co-dominance-and-incomplete-dominance}}

a co-dominant colour combination can look like an intermediate if it is
the combination of two pigments, however in the case where one of the
phenotypes is white then a intermediate can only be a result of
incomplete dominance

\hypertarget{sex-linked-traits}{%
\paragraph{Sex linked traits}\label{sex-linked-traits}}

\hypertarget{sex-limited-traits}{%
\subparagraph{Sex limited traits,}\label{sex-limited-traits}}

\hypertarget{important-examples}{%
\subsubsection{Important examples}\label{important-examples}}

BG 2032

\hypertarget{section}{%
\subsection{13/02/2108}\label{section}}

\hypertarget{phenylketonuria}{%
\subsubsection{Phenylketonuria}\label{phenylketonuria}}

single mutation can affect brain development however if the correct diet
is followed then all detrimental health consequences can be ignored.

\hypertarget{diabetes-type-ii}{%
\subsubsection{diabetes type II}\label{diabetes-type-ii}}

lifestyle disease, the genetic component is estimated to change the
likelihood of developing the disease by 2\%, but the main causative
factors of the disease relate to diet.

\hypertarget{lactose-intolerance.}{%
\subsubsection{Lactose intolerance.}\label{lactose-intolerance.}}

Individuals with lactose intolerance cant digest lactose, however the
bacteria in their gut van which leads to the build up of gas. this can
be very painful for the person in question.In humans lactose intolerance
would naturally occur after weening however as cow milk is a staple part
of most human diabetes the genes controlling lactase production are
never switched off and the ability to digest lactose is not normally
lost.

\hypertarget{himalayan-phenotype.}{%
\subsubsection{Himalayan Phenotype.}\label{himalayan-phenotype.}}

\hypertarget{cancer}{%
\subsubsection{Cancer}\label{cancer}}

\hypertarget{basics}{%
\paragraph{basics}\label{basics}}

A cancer is a group of cells whose proliferation is uncontrolled, and
which can spread to other locations in the body which are normally
populated bu other cell types.

\hypertarget{benigin-tumours}{%
\subparagraph{benigin tumours}\label{benigin-tumours}}

grow but do not spread.

\hypertarget{malignant-tumours}{%
\subparagraph{malignant tumours}\label{malignant-tumours}}

over proliferate and invade other body tissues/ areas in the body.

\hypertarget{types-of-cancers}{%
\paragraph{types of cancers}\label{types-of-cancers}}

\hypertarget{carcinomas}{%
\subparagraph{Carcinomas}\label{carcinomas}}

derived form epithelial cells

\hypertarget{sarcomas}{%
\subparagraph{Sarcomas}\label{sarcomas}}

derived from connective or muscle tissues. Known as Osteosarcoma in the
case of bone tissue and Kaposi sarcoma in the case of soft connective
tissue.

\hypertarget{lymphoma-and-leukaemias}{%
\subparagraph{Lymphoma and leukaemias}\label{lymphoma-and-leukaemias}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Cancers of the hematopoietic system.
\item
  Lymphoma (solid tumour)
\item
  Leukaemia excess of circulating immature blood cell precursors.
  \#\#\#\#\# Cancer of the nervous system brain and central nervous
  system.
\end{enumerate}

NOTE: most common in children

\hypertarget{germ-line-cancer.}{%
\subparagraph{Germ-line cancer.}\label{germ-line-cancer.}}

\hypertarget{cancers-of-breast-prostate-lung-pancreas-and-colon.}{%
\subparagraph{Cancers of breast, prostate, lung, pancreas, and
colon.}\label{cancers-of-breast-prostate-lung-pancreas-and-colon.}}

\hypertarget{causes-of-cancer.}{%
\subsubsection{Causes of cancer.}\label{causes-of-cancer.}}

In most cases cancer is not inherited. In breast cancer/ovarian cancer
BS1, BS2 play some role in genetic disposition, perhaps 5-10\%

\hypertarget{hereditary.}{%
\subsubsection{hereditary.}\label{hereditary.}}

Cancer normally sets in during old age as cells must accumulate a series
of a specific set of notations, (exactly which set of mutations
accumulated lead to cancer is highly individual)

\hypertarget{environmental-factors}{%
\paragraph{Environmental factors}\label{environmental-factors}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  UV
\item
  X rays
\item
  Alcohol
\item
  Overcooked food (hetercyclic amines,polycyclic aromatic
  hydrocarbons-meat. acrylamide- potatoes
\item
  Azo dyes
\item
  tartrazine (food colourants)
\item
  Nitrate cured foods
\item
  Pesticides
\end{enumerate}

\hypertarget{cancer-initiators}{%
\paragraph{Cancer initiators}\label{cancer-initiators}}

agents which cause DNA damage (mutagens). these factors may be chemical,
biological (such as HPV- human papilloma virus cause latent genetic
damage- Cancer predisposition.

\hypertarget{cancer-promoters}{%
\paragraph{Cancer Promoters}\label{cancer-promoters}}

Promotes excessive proliferation (does not directly damage DNA).

\hypertarget{examples}{%
\subparagraph{Examples}\label{examples}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  wounding
\item
  phorbol esters
\item
  HRT/ oestrogen (breast cancer)
\item
  hepatitis B (promotes stomach cancer)
\item
  HIV (Kaposi sarcoma, this disease is always present but only manifests
  itself in immunocompromised individuals)
\end{enumerate}

\hypertarget{stages-of-cancer-development}{%
\paragraph{stages of cancer
development}\label{stages-of-cancer-development}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  normal cells (mutation occcurs)
\item
  increased ability to proliferate (mutation)
\item
  1-2mm tumour of rapidly proliferating cells which do not undergo
  apoptosis of differentiation. (mutation)
\item
  vascularized growing tumour (mutation)
\item
  large tumour capable of invading near by tissue (metastasis)
\end{enumerate}

NOTE: cancer results from a series of somatic mutations, affecting the
same cell, and different cancers are genetically heterogeneous.

\hypertarget{genetic-instability-in-cancers.}{%
\paragraph{Genetic instability in
cancers.}\label{genetic-instability-in-cancers.}}

deficient local DNA repair leads to the accumulation of point mutations
increased chromosomal instability and gross genome abnormalities.

\hypertarget{summary-of-key-properties}{%
\paragraph{summary of key properties}\label{summary-of-key-properties}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  disregards ex/in growth regulation signals.
\item
  avoid apoptosis, differentiation and replicative senescence
\item
  genetically unstable.
\item
  invasive
\item
  metastatic (survive and proliferate in foreign) sites.
\end{enumerate}

\hypertarget{penetrance}{%
\paragraph{Penetrance}\label{penetrance}}

percentage of the population who demonstrate at least some degree fo
phenotypic expression.

\hypertarget{expressivity}{%
\paragraph{Expressivity}\label{expressivity}}

reflects the range of expressions of the gene/allele present in the
population.

\hypertarget{genetic-complementation}{%
\paragraph{Genetic complementation}\label{genetic-complementation}}

two different mutations in heterozygous condition affecting the same
protein/ pathway can compliment each other to cause a novel phenotypic
effect.

\hypertarget{examples-1}{%
\subparagraph{examples}\label{examples-1}}

rare form of albinism: both mother and father were normally pigmented
and there families had minimal instances of albinism. however both
children where albino' s . Analysis found that this albinism stemmed not
from those genes and mutation most commonly associated with albinism but
rather from a combination of recessives (heterozygous) from both father
and mother (mutations in TYRP1 gene), which on their own would have
little or no effect on the pigment production pathway, but when combined
reduced pigment production significantly.

NOTE: Rufus albinism leads to a phenotype with reddish hair, lighter
skin, and blue grey eyes.

\hypertarget{forked-line-probability-method.}{%
\paragraph{Forked line probability
method.}\label{forked-line-probability-method.}}

\#\#chromosome level (revise mitosis and meosis) copy slides chromosome
basic structure \#\#\#\# chromosome groups

\hypertarget{meta-centric}{%
\subparagraph{meta-centric}\label{meta-centric}}

centromere is half way up the length of the chromosome

\hypertarget{submetacentric}{%
\subparagraph{submetacentric}\label{submetacentric}}

centromere is more to the one side of the chromosome than the other,

\hypertarget{acrocentric}{%
\subparagraph{Acrocentric}\label{acrocentric}}

the centromere is very far to the one side of the chromosome, with a
long arm containing most of the genes and a short arm containing
predominantly temolmeric DNA.

\hypertarget{telocentric}{%
\subparagraph{Telocentric}\label{telocentric}}

humans do not posses any telocentric chromosomes, but certain insect of
crustation species do.

\hypertarget{holocentric}{%
\subparagraph{Holocentric}\label{holocentric}}

Centromere like structure exist along the entire length of the
chromosome. this may decrease the chances of faulty division/
segregation.

\hypertarget{chromosome-banding}{%
\subparagraph{chromosome banding}\label{chromosome-banding}}

bands were named and used to locate specific genes. the banding patterns
are due to uneven DNA densities in the coiled structure of the
chromosome.

\hypertarget{size-and-shape-of-different-chromosomes.}{%
\subparagraph{Size and shape of different
chromosomes.}\label{size-and-shape-of-different-chromosomes.}}

\hypertarget{chromosome-level-mutations}{%
\subsubsection{Chromosome level
mutations}\label{chromosome-level-mutations}}

\hypertarget{aneuploidy-spelling}{%
\paragraph{Aneuploidy (Spelling)}\label{aneuploidy-spelling}}

Each cell is has at least one extra chromosome or is missing at least
one chromosome

\hypertarget{trisomy}{%
\subparagraph{Trisomy}\label{trisomy}}

when an individual inherits three copies of a particular chromosome. In
Humans only three of all the possible trisomies are viable, (as in
individuals with these mutations will still be born and not terminated
during pregnancy)

of these three (Trisomy 13, 18, and 21) only individuals with trisomy 21
can survive past the first few years of childhood. Individuals with
Trisomy 21 have down syndrome.

Down syndrome (Trisomy 21)

Frequency: 1/1000 Effects 1. mental retardation 2. short stature 3.
heart disease 4. shortened life span.

Edward Syndrome (Trisomy 18)

Effects: 1. kidney and intestinal malformation 3. heart defects. 4.
mental retardation

NOTE: only 8\% survival past the first year.

Patua syndrome (Trisomy 13 )

Effects: 1. Kidney and hear defects 2. polidactily (too many digits) 3.
Nervous system abnormalities 4. Death within the 1st year.

NOTE: the smaller relative size of chromosome 21, meaning that it
contains less important genes may be related to its increased viability
.

\hypertarget{monosomy}{%
\subparagraph{Monosomy}\label{monosomy}}

when an individual inherits only one copy of a particular chromosome.

\hypertarget{nullisomy}{%
\subparagraph{Nullisomy}\label{nullisomy}}

having no copies of a particular chromosome.

\hypertarget{polyploidy}{%
\paragraph{Polyploidy}\label{polyploidy}}

\hypertarget{structural-rearrangements}{%
\paragraph{Structural rearrangements}\label{structural-rearrangements}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion
\item
  duplication
\item
  translocation
\item
  inversion
\item
  fission
\item
  fusion
\end{enumerate}

\hypertarget{non-disjuncture}{%
\paragraph{non disjuncture}\label{non-disjuncture}}

(review mitosis and meiosis) If non disjuncture happens in meiosis it is
generally worse than if it happens in mitosis as it will affect the
whole organism.

\hypertarget{down-syndrome}{%
\subparagraph{Down Syndrome}\label{down-syndrome}}

occurs in about 1 of a thousand live births. originates from non
disjunction in the egg cell rather than the sperm cell because the
arrested development of the egg cells leads to decay of the separating
and marking proteins as well as the spindle fibre leaving the whole
process open to more error. 8\% of individuals survive for one year ?

\hypertarget{turner-syndrome-monosomy-x}{%
\subparagraph{turner Syndrome monosomy
X}\label{turner-syndrome-monosomy-x}}

Frequency 1/20 000 (only occurs in females ) Effects: 1. webbed neck 2.
short stature 3. underdeveloped ovaries and lack of secondary sex
characteristics.

\hypertarget{klinefelter-syndrome-xxy}{%
\subparagraph{Klinefelter syndrome XXY}\label{klinefelter-syndrome-xxy}}

Too many X's ,XXY , XXXY, XXXXY, XXXXXY. (XXYY) Effects: 1. Decrease in
testosterone levels 2. the more X's the higher the chance of brain
damage.

\hypertarget{xyy}{%
\subparagraph{XYY}\label{xyy}}

XYY has the phenotype of a normal make. men with 2 Y chromosomes tend to
be tall as there is a cumulative effect adding to height. (they may also
exhibit increased aggression but this in unclear)

\hypertarget{xxx}{%
\subparagraph{XXX}\label{xxx}}

Frequency: 1/1000 Effects: 1. Normal female with normal fertility.

NOTE: this conditions is seldom diagnosed.

\hypertarget{polyploidy-1}{%
\paragraph{Polyploidy}\label{polyploidy-1}}

Three or more complete sets of chromosomes present in somatic cells.
\#\#\#\# 2n-20n caused by non reduced gamete fomration, which is very
rare in males and non viable.

most often flowering plants are polyploid, they are even specifically
breed to have more chromosome sets as this usually increases fruit and
flower size.

\hypertarget{examples-2}{%
\subparagraph{Examples}\label{examples-2}}

water melon must be bread from a tetraploid(?) and a diploid to get a
tripliod infertile plant.

Kiwi fruit 12-16 copies

Strawberries 4 copies (tetraploid)

Frogs \emph{Xenopis levis} 4n \emph{Xenopus tropicalis} 2n

NOTE: plants are better adapted to polyploid because: 1. they are not as
confined to a set physical form so different in growth and development
genes operation levels are not so important 2. they can reproduce
vegetatively so polyploid individuals aren't as severely evolutionarily
disadvantaged. 3. less precise sperm targeting is necessary.

\hypertarget{deletions}{%
\paragraph{Deletions}\label{deletions}}

deletions often occur by chromosome breakage during the cell cycle.
These deletions may be terminal of interstitial. If the centromere is
lost, the entire chromosome will be lost.

\hypertarget{piece-of-chrome}{%
\subparagraph{Piece of Chrome}\label{piece-of-chrome}}

sometimes multiple genes are lost when both strands break and a part of
the chromosome is permanently lost. A specific deletion on chromosome 5.
(where the entire p arm is deleted leads to a serious syndrome)

\hypertarget{cri-du-chat-5p-}{%
\subparagraph{Cri du chat (5p-)}\label{cri-du-chat-5p-}}

Frequency: 1/20000-50000 Effects: 1. Intellectual disability, 2. low
muscle tone 3. microcephaly 4. distinctive facial features. 5. normal
life expectancy

\hypertarget{duplications}{%
\paragraph{Duplications}\label{duplications}}

Cause: retro transposition, or non-allelic homologous recombination in
repeat rich areas.

A small region of a chromosome is repeated, and placed next to the
existing copy (in tandem). duplications tend to happen in high
repetitive sequences of DNA. These repeated regions mean that the
chromosome will note line up very well with its homologous pair.

passed on by unequal order, and may or may not have a phenotypic effect.

\hypertarget{huntington-purea}{%
\paragraph{Huntington Purea}\label{huntington-purea}}

Repeats alter gene function, which can lead to brain damage. Repeats may
have a positive effect because they allow one copy of a gene to evolve
independently, specialise in a different function while the original
function is still conserved by the other copy of the gene.

\hypertarget{inversions.}{%
\paragraph{Inversions.}\label{inversions.}}

Cause: non-allelic homologous recombination Effects: 1. pairing in
meiosis occurs via looping 2. cross over in the looped region can cause
major deletions or duplications.

\hypertarget{translocation}{%
\paragraph{Translocation}\label{translocation}}

Effects: 1. novel chromosomes are generated 2. individuals are usually
infertile. 3. can result in aneuploidy (such as 14q21q inherited down
syndrome) 4.

\hypertarget{terminal}{%
\subparagraph{terminal}\label{terminal}}

A piece breaks off a chromosome and sticks to the end of a
non-homologous chromosome.

\hypertarget{reciprocal}{%
\subparagraph{reciprocal}\label{reciprocal}}

Exchange of pieces between non-homologous chromosomes \#\#\#\#
Robertsonian

two Acrocentric chromosomes lose their short arms and get stuck
together.

\hypertarget{fission}{%
\paragraph{Fission}\label{fission}}

deletion (loss of acentric piece)'

\hypertarget{fusion}{%
\paragraph{Fusion}\label{fusion}}

bicentric chromosomes (breakage in mitosis)

\hypertarget{breakage-fusion-bridge-cycles-in-cancer.}{%
\subparagraph{breakage fusion bridge cycles in
cancer.}\label{breakage-fusion-bridge-cycles-in-cancer.}}

Copy image. 1. end of a chromosome breaks off (during meiosis) 2. one
daughter cell inherits the chromosome lacking a telomere. 3. the new
cell enters S phase and replicates its DNA (it now has two sticky DNA
ends right next to each other. ) 4. loose DNA strands bind to form
continuous DNA loop. 5. fused sister chromatids are pulled apart during
mitosis, causing breakage at a new sight 6. one daughter cell inherits
chromosome with duplicated genes and again a free sticky end. this cycle
can then begin again leading to even further chromosome distortion.

\hypertarget{imprinting}{%
\paragraph{Imprinting}\label{imprinting}}

\hypertarget{hinny-vs-mule}{%
\subparagraph{Hinny vs mule}\label{hinny-vs-mule}}

Mule

horse mother donkey father hardy and obedient

Hinny

donkey mother, horse father. temperamental, untrainable.

\hypertarget{intersistial-deletion-of-chromosome-15}{%
\subparagraph{Intersistial deletion of chromosome
15}\label{intersistial-deletion-of-chromosome-15}}

Prader -Willi syndrome

Cause: deletion from father, ZNF127 and IPW inactivated on the maternal
copy.\\
effects: 1. mental retardation 2. obesity 3. diminished growth.

Angelman syndrome

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion from mother: UBE3A inactivated on paternal copy.\\
\item
  mental retardation
\item
  epilepsy
\item
  lack of motor development.
\end{enumerate}

\hypertarget{x-inactivation}{%
\paragraph{X inactivation}\label{x-inactivation}}

the inactivation of one of two X chromosomes in females via DNA
methylation.

NOTE: which X chromosome is inactivated varies randomly from cell to
cell.this inactivation process turns the inactive X into a bar body,
(highly methylated and rolled up in histone proteins).

\hypertarget{frog-colour}{%
\paragraph{Frog colour}\label{frog-colour}}

\hypertarget{details}{%
\subparagraph{details}\label{details}}

the skin colour of a certain from is controlled by one allele. There are
three different phenotypes observed, these are red, blue, and purple.

\hypertarget{explanation}{%
\subparagraph{explanation}\label{explanation}}

This is an example of co-dominance as the purple colour comes from a
combination of the red and blue pigments. that is both red and blue
pigments have been produced. We can conclude that the purple from is
heterozygous, the blue frog is homozygous blue and the red frog is
homozygous red.

\hypertarget{fruit-flies-eye-colour.}{%
\paragraph{Fruit flies eye colour.}\label{fruit-flies-eye-colour.}}

\hypertarget{details.}{%
\subparagraph{details.}\label{details.}}

the colour of fruit flies eyes is determined by two gene loci. a fly
with dominant alleles at the red locus will be red eyed, while a fly
with dominant alleles at the brown locus will be brown eyed. However a
fly with both dominant brown and cinnabar will be white eyed.

\hypertarget{explanation-1}{%
\subparagraph{explanation}\label{explanation-1}}

the white eyed phenotype is due to a form of inhibition similar to
epistasis stops either of the pigments form being produced if both
pigment genes are present.

\hypertarget{rabbit-skin-colour}{%
\paragraph{Rabbit Skin Colour}\label{rabbit-skin-colour}}

\hypertarget{details-1}{%
\subparagraph{details}\label{details-1}}

the skin colour of rabbits is determined by one gene locus, which has
four possible Alleles. All alleles code for the enzyme tyrosinase which
acts to convert tyrosine into dopamine a critical first set in the
production of hair pigment.

\#1 non functional

The most recessive Allele codes for a nonfunctioning version of
tyrosinase leading to albino rabbit.

\#2 heat sensitive

The next most recessive allele codes for a heat inactivated version of
tyrosine meaning that the final colour of the rabbits fur will depend on
the temperature conditions it was raised in, and also that within a
certain temperature range there will be a distinct colour difference
between the warmer and colder parts of the animal.

\#3 localised functionality

The next most recessive allele codes for a version of tyrosinase does
not fold quite ``correctly'' and as such does not glycosalate in the
same way as the dominant version. melanocytes which produce the grey
pigment will still accept this tyrosinase into their cytoplasm and as
such grey pigment will be produced. however as intracellular transport
is affected the tyrosinase either cannot reach or cannot enter
melanocytes capable of producing the brown pigments and so they rabbits
are full grey

\#4 fully functional

the most dominant allele codes for a fully functional form of tyrosinase
and the combination of grey and orange pigments produced gives the
rabbits a full brown coat.

\hypertarget{human-blood-types.}{%
\paragraph{human Blood types.}\label{human-blood-types.}}

\hypertarget{details-2}{%
\paragraph{details}\label{details-2}}

human blood types are determined by one gene locus with three alleles.
allele O is recessive, and codes for no functioning protein. allele A
and B are co dominant and each code for different sugars which are
attached to the outside of red blood cells.

\hypertarget{bombay-phenotype}{%
\paragraph{Bombay phenotype}\label{bombay-phenotype}}

\hypertarget{details-3}{%
\subparagraph{details}\label{details-3}}

parents A and AB, child O, mates with A, results B, AB and O.

\hypertarget{explanation-2}{%
\subparagraph{explanation}\label{explanation-2}}

A and B sugars are attached to red blood cells by appending them onto
the end of glycoproteins called Substance H which are already attached
to the cell. If a individual is homozygous recessive for substance H
genes, none will be produced and so even if the have A or B alleles they
will have a functionally O phenotype.

\#\#\#Alleles and genes A gene is a collection/ group of alleles which
could occur the same locus and which are associated with the same
traits.

\#\#\#lethal alleles alleles which are incompatible with the organisms
continued survival

\hypertarget{dominance-1}{%
\paragraph{dominance}\label{dominance-1}}

most lethal alleles are recessive because dominant lethal alleles
produce non functioning proteins only so any individual with even one of
the dominant allele would be unlikely to survive to reproductive age,
and so the Allele would very quickly be bread out of the population.

\hypertarget{sex-chromosomes}{%
\paragraph{sex chromosomes}\label{sex-chromosomes}}

lethal alleles, or alleles linked with low fertility located on the Y
chromosome in humans are very rare. disease related to homozygous XX are
more likely to be suffered by woman, as they will have one working copy
of the given gene so they can survive but one defective copy which leads
to the illness. males on the other hand will either survive and be
entirely healthy with a working copy or die, if they are even born if
they have only a defective copy.

\hypertarget{lethal-dominants}{%
\subparagraph{lethal dominants}\label{lethal-dominants}}

lethal dominants may occur in one of two cases. First, if the onset of
the disease/ defect related to the Allele only occurs after reproductive
age. Secondly if a De Novo (random, once off) DNA mutation occurs in a
given individual. lethal dominants can also result from an Allele which
interferes with the formation of aggregates of the protein which the
gene codes for, or receptors of that protein,or finally if it has a
detrimental effect on another protein or gene (such as the yellow mice)

\hypertarget{examples-3}{%
\paragraph{examples}\label{examples-3}}

\hypertarget{yellow-mice.}{%
\subparagraph{Yellow Mice.}\label{yellow-mice.}}

\hypertarget{details-4}{%
\subparagraph{details}\label{details-4}}

Y allele codes for the yellow coloured pigment. y codes for other
pigment/no pigment. Y is a lethal Dominant(?) Allele. As a YY individual
with not develop bast the early stages of implantation into the uterine
wall.

\hypertarget{moleculargenetic-explanation}{%
\subparagraph{molecular/genetic
explanation}\label{moleculargenetic-explanation}}

When the Y allele is present at the gene loci it causes a deletion in
which, its own gene promoter, as well as the coding sequence of the gene
upstream of it(the MERC gene necessary in RNA processing) are removed.
the net result being that proteins relating to the gene upstream are no
longer removed, and the pigment gene is now promoter by the MERC
promoter instead of its own promoter

\hypertarget{result-and-phenotype}{%
\subparagraph{Result and Phenotype}\label{result-and-phenotype}}

the result is that Yy individual still produce enough RNA processing
proteins to survive/ avoid miscarriage,but they are more susceptible to
diseases such as obesity and cancer. Also their yellow pigment gene is
promoted heavily during early development leading to yellow coloured
fur. YY miscarriages early in pregnancy as it cant produce necessary RNA
processing proteins. yy is health individual without yellow colour.
birth rations will be skewed by the fact that YY individuals are not
actually born. NOTE: this is an example of incomplete dominance of Y
with respect to MERC.

\hypertarget{effective-of-the-environment-on-genetic-expression.}{%
\subsubsection{Effective Of the environment on genetic
expression.}\label{effective-of-the-environment-on-genetic-expression.}}

\hypertarget{examples-4}{%
\paragraph{examples}\label{examples-4}}

Myopia(short sightedness):

Populations where book learning is not inforced/encouraged from an early
age do not show high rates of myopia even though the genetic dispostion
between the two populations is very similar.

\hypertarget{phenylketonuria-1}{%
\subsubsection{Phenylketonuria}\label{phenylketonuria-1}}

single mutation can affect brain development however if the correct diet
is followed then all detrimental health consequences can be ignored.

\hypertarget{diabetes-type-ii-1}{%
\subsubsection{Diabetes type II}\label{diabetes-type-ii-1}}

lifestyle disease, the genetic component is estimated to change the
likelihood of developing the disease by 2\%, but the main causative
factors of the disease relate to diet.

\hypertarget{lactose-intolerance.-1}{%
\subsubsection{Lactose intolerance.}\label{lactose-intolerance.-1}}

Individuals with lactose intolerance cant digest lactose, however the
bacteria in their gut van which leads to the build up of gas. this can
be very painful for the person in question.In humans lactose intolerance
would naturally occur after weening however as cow milk is a staple part
of most human diabetes the genes controlling lactase production are
never switched off and the ability to digest lactose is not normally
lost.

\hypertarget{himalayan-phenotype.-1}{%
\subsubsection{Himalayan Phenotype.}\label{himalayan-phenotype.-1}}

\hypertarget{cancer-1}{%
\subsubsection{Cancer}\label{cancer-1}}

\hypertarget{basics-1}{%
\paragraph{basics}\label{basics-1}}

A cancer is a group of cells whose proliferation is uncontrolled, and
which can spread to other locations in the body which are normally
populated bu other cell types.

\hypertarget{benigin-tumours-1}{%
\subparagraph{benigin tumours}\label{benigin-tumours-1}}

grow but do not spread.

\hypertarget{malignant-tumours-1}{%
\subparagraph{malignant tumours}\label{malignant-tumours-1}}

over proliferate and invade other body tissues/ areas in the body.

\hypertarget{types-of-cancers-1}{%
\paragraph{types of cancers}\label{types-of-cancers-1}}

\hypertarget{carcinomas-1}{%
\subparagraph{Carcinomas}\label{carcinomas-1}}

derived form epithelial cells

\hypertarget{sarcomas-1}{%
\subparagraph{Sarcomas}\label{sarcomas-1}}

derived from connective or muscle tissues. Known as Osteosarcoma in the
case of bone tissue and Kaposi sarcoma in the case of soft connective
tissue.

\hypertarget{lymphoma-and-leukaemias-1}{%
\subparagraph{Lymphoma and leukaemias}\label{lymphoma-and-leukaemias-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Cancers of the hematopoietic system.
\item
  Lymphoma (solid tumour)
\item
  Leukaemia excess of circulating immature blood cell precursors.
  \#\#\#\#\# Cancer of the nervous system brain and central nervous
  system.
\end{enumerate}

NOTE: most common in children

\hypertarget{germ-line-cancer.-1}{%
\subparagraph{Germ-line cancer.}\label{germ-line-cancer.-1}}

\hypertarget{cancers-of-breast-prostate-lung-pancreas-and-colon.-1}{%
\subparagraph{Cancers of breast, prostate, lung, pancreas, and
colon.}\label{cancers-of-breast-prostate-lung-pancreas-and-colon.-1}}

\hypertarget{causes-of-cancer.-1}{%
\subsubsection{Causes of cancer.}\label{causes-of-cancer.-1}}

In most cases cancer is not inherited. In breast cancer/ovarian cancer
BS1, BS2 play some role in genetic disposition, perhaps 5-10\%

\hypertarget{hereditary.-1}{%
\subsubsection{Hereditary.}\label{hereditary.-1}}

Cancer normally sets in during old age as cells must accumulate a series
of a specific set of notations, (exactly which set of mutations
accumulated lead to cancer is highly individual) \#\#\#\# Environmental
factors 1. UV 2. X rays 3. Alcohol 4. Overcooked food (hetercyclic
amines,polycyclic aromatic hydrocarbons-meat. acrylamide- potatoes 5.
Azo dyes 6. tartrazine (food colourants) 7. Nitrate cured foods 8.
Pesticides

\hypertarget{cancer-initiators-1}{%
\paragraph{Cancer initiators}\label{cancer-initiators-1}}

agents which cause DNA damage (mutagens). these factors may be chemical,
biological (such as HPV- human papilloma virus cause latent genetic
damage- Cancer predisposition.

\hypertarget{cancer-promoters-1}{%
\paragraph{Cancer Promoters}\label{cancer-promoters-1}}

Promotes excessive proliferation (does not directly damage DNA).

\hypertarget{examples-5}{%
\subparagraph{Examples}\label{examples-5}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  wounding
\item
  phorbol esters
\item
  HRT/ oestrogen (breast cancer)
\item
  hepatitis B (promotes stomach cancer)
\item
  HIV (Kaposi sarcoma, this disease is always present but only manifests
  itself in immunocompromised individuals)
\end{enumerate}

\hypertarget{stages-of-cancer-development-1}{%
\paragraph{stages of cancer
development}\label{stages-of-cancer-development-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  normal cells (mutation occcurs)
\item
  increased ability to proliferate (mutation)
\item
  1-2mm tumour of rapidly proliferating cells which do not undergo
  apoptosis of differentiation. (mutation)
\item
  vascularized growing tumour (mutation)
\item
  large tumour capable of invading near by tissue (metastasis)
\end{enumerate}

NOTE: cancer results from a series of somatic mutations, affecting the
same cell, and different cancers are genetically heterogeneous.

\hypertarget{genetic-instability-in-cancers.-1}{%
\paragraph{Genetic instability in
cancers.}\label{genetic-instability-in-cancers.-1}}

deficient local DNA repair leads to the accumulation of point mutations
increased chromosomal instability and gross genome abnormalities.

\hypertarget{summary-of-key-properties-1}{%
\paragraph{summary of key
properties}\label{summary-of-key-properties-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  disregards ex/in growth regulation signals.
\item
  avoid apoptosis, differentiation and replicative senescence
\item
  genetically unstable.
\item
  invasive
\item
  metastatic (survive and proliferate in foreign) sites.
\end{enumerate}

\hypertarget{penetrance-1}{%
\paragraph{Penetrance}\label{penetrance-1}}

percentage of the population who demonstrate at least some degree fo
phenotypic expression.

\hypertarget{expressivity-1}{%
\paragraph{Expressivity}\label{expressivity-1}}

reflects the range of expressions of the gene/allele present in the
population.

\hypertarget{genetic-complementation-1}{%
\paragraph{Genetic complementation}\label{genetic-complementation-1}}

two different mutations in heterozygous condition affecting the same
protein/ pathway can compliment each other to cause a novel phenotypic
effect.

\hypertarget{examples-6}{%
\subparagraph{examples}\label{examples-6}}

rare form of albinism: both mother and father were normally pigmented
and there families had minimal instances of albinism. however both
children where albino' s . Analysis found that this albinism stemmed not
from those genes and mutation most commonly associated with albinism but
rather from a combination of recessives (heterozygous) from both father
and mother (mutations in TYRP1 gene), which on their own would have
little or no effect on the pigment production pathway, but when combined
reduced pigment production significantly.

NOTE: Rufus albinism leads to a phenotype with reddish hair, lighter
skin, and blue grey eyes.

\hypertarget{forked-line-probability-method.-1}{%
\paragraph{Forked line probability
method.}\label{forked-line-probability-method.-1}}

\#\#Chromosome level (revise mitosis and meosis) copy slides chromosome
basic structure \#\#\#\# chromosome groups

\hypertarget{meta-centric-1}{%
\subparagraph{meta-centric}\label{meta-centric-1}}

centromere is half way up the length of the chromosome

\hypertarget{submetacentric-1}{%
\subparagraph{submetacentric}\label{submetacentric-1}}

centromere is more to the one side of the chromosome than the other,

\hypertarget{acrocentric-1}{%
\subparagraph{Acrocentric}\label{acrocentric-1}}

the centromere is very far to the one side of the chromosome, with a
long arm containing most of the genes and a short arm containing
predominantly temolmeric DNA.

\hypertarget{telocentric-1}{%
\subparagraph{Telocentric}\label{telocentric-1}}

humans do not posses any telocentric chromosomes, but certain insect of
crustation species do.

\hypertarget{holocentric-1}{%
\subparagraph{Holocentric}\label{holocentric-1}}

Centromere like structure exist along the entire length of the
chromosome. this may decrease the chances of faulty division/
segregation.

\hypertarget{chromosome-banding-1}{%
\subparagraph{chromosome banding}\label{chromosome-banding-1}}

bands were named and used to locate specific genes. the banding patterns
are due to uneven DNA densities in the coiled structure of the
chromosome.

\hypertarget{size-and-shape-of-different-chromosomes.-1}{%
\subparagraph{Size and shape of different
chromosomes.}\label{size-and-shape-of-different-chromosomes.-1}}

\hypertarget{chromosome-level-mutations-1}{%
\subsubsection{Chromosome level
mutations}\label{chromosome-level-mutations-1}}

\hypertarget{aneuploidy-spelling-1}{%
\paragraph{Aneuploidy (Spelling)}\label{aneuploidy-spelling-1}}

Each cell is has at least one extra chromosome or is missing at least
one chromosome

\hypertarget{trisomy-1}{%
\subparagraph{Trisomy}\label{trisomy-1}}

when an individual inherits three copies of a particular chromosome. In
Humans only three of all the possible trisomies are viable, (as in
individuals with these mutations will still be born and not terminated
during pregnancy)

of these three (Trisomy 13, 18, and 21) only individuals with trisomy 21
can survive past the first few years of childhood. Individuals with
Trisomy 21 have down syndrome.

Down syndrome (Trisomy 21)

Frequency: 1/1000 Effects 1. mental retardation 2. short stature 3.
heart disease 4. shortened life span.

Edward Syndrome (Trisomy 18)

Effects: 1. kidney and intestinal malformation 3. heart defects. 4.
mental retardation

NOTE: only 8\% survival past the first year.

Patua syndrome (Trisomy 13 )

Effects: 1. Kidney and hear defects 2. polidactily (too many digits) 3.
Nervous system abnormalities 4. Death within the 1st year.

NOTE: the smaller relative size of chromosome 21, meaning that it
contains less important genes may be related to its increased viability
.

\hypertarget{monosomy-1}{%
\subparagraph{Monosomy}\label{monosomy-1}}

when an individual inherits only one copy of a particular chromosome.

\hypertarget{nullisomy-1}{%
\subparagraph{Nullisomy}\label{nullisomy-1}}

having no copies of a particular chromosome.

\hypertarget{polyploidy-2}{%
\paragraph{Polyploidy}\label{polyploidy-2}}

\hypertarget{structural-rearrangements-1}{%
\paragraph{Structural
rearrangements}\label{structural-rearrangements-1}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion
\item
  duplication
\item
  translocation
\item
  inversion
\item
  fission
\item
  fusion
\end{enumerate}

\hypertarget{non-disjuncture-1}{%
\paragraph{non disjuncture}\label{non-disjuncture-1}}

(review mitosis and meiosis) If non disjuncture happens in meiosis it is
generally worse than if it happens in mitosis as it will affect the
whole organism.

\hypertarget{down-syndrome-1}{%
\subparagraph{Down Syndrome}\label{down-syndrome-1}}

occurs in about 1 of a thousand live births. originates from non
disjunction in the egg cell rather than the sperm cell because the
arrested development of the egg cells leads to decay of the separating
and marking proteins as well as the spindle fibre leaving the whole
process open to more error. 8\% of individuals survive for one year ?

\hypertarget{turner-syndrome-monosomy-x-1}{%
\subparagraph{turner Syndrome monosomy
X}\label{turner-syndrome-monosomy-x-1}}

Frequency 1/20 000 (only occurs in females ) Effects: 1. webbed neck 2.
short stature 3. underdeveloped ovaries and lack of secondary sex
characteristics.

\hypertarget{klinefelter-syndrome-xxy-1}{%
\subparagraph{Klinefelter syndrome
XXY}\label{klinefelter-syndrome-xxy-1}}

Too many X's ,XXY , XXXY, XXXXY, XXXXXY. (XXYY) Effects: 1. Decrease in
testosterone levels 2. the more X's the higher the chance of brain
damage.

\hypertarget{xyy-1}{%
\subparagraph{XYY}\label{xyy-1}}

XYY has the phenotype of a normal make. men with 2 Y chromosomes tend to
be tall as there is a cumulative effect adding to height. (they may also
exhibit increased aggression but this in unclear)

\hypertarget{xxx-1}{%
\subparagraph{XXX}\label{xxx-1}}

Frequency: 1/1000 Effects: 1. Normal female with normal fertility.

NOTE: this conditions is seldom diagnosed.

\hypertarget{polyploidy-3}{%
\paragraph{Polyploidy}\label{polyploidy-3}}

Three or more complete sets of chromosomes present in somatic cells.
\#\#\#\# 2n-20n caused by non reduced gamete fomration, which is very
rare in males and non viable.

most often flowering plants are polyploid, they are even specifically
breed to have more chromosome sets as this usually increases fruit and
flower size.

\hypertarget{examples-7}{%
\subparagraph{Examples}\label{examples-7}}

water melon must be bread from a tetraploid(?) and a diploid to get a
tripliod infertile plant.

Kiwi fruit 12-16 copies

Strawberries 4 copies (tetraploid)

Frogs \emph{Xenopis levis} 4n \emph{Xenopus tropicalis} 2n

NOTE: plants are better adapted to polyploid because: 1. they are not as
confined to a set physical form so different in growth and development
genes operation levels are not so important 2. they can reproduce
vegetatively so polyploid individuals aren't as severely evolutionarily
disadvantaged. 3. less precise sperm targeting is necessary.

\hypertarget{deletions-1}{%
\paragraph{Deletions}\label{deletions-1}}

deletions often occur by chromosome breakage during the cell cycle.
These deletions may be terminal of interstitial. If the centromere is
lost, the entire chromosome will be lost.

\hypertarget{piece-of-chrome-1}{%
\subparagraph{Piece of Chrome}\label{piece-of-chrome-1}}

sometimes multiple genes are lost when both strands break and a part of
the chromosome is permanently lost. A specific deletion on chromosome 5.
(where the entire p arm is deleted leads to a serious syndrome)

\hypertarget{cri-du-chat-5p--1}{%
\subparagraph{Cri du chat (5p-)}\label{cri-du-chat-5p--1}}

Frequency: 1/20000-50000 Effects: 1. Intellectual disability, 2. low
muscle tone 3. microcephaly 4. distinctive facial features. 5. normal
life expectancy

\hypertarget{duplications-1}{%
\paragraph{Duplications}\label{duplications-1}}

Cause: retro transposition, or non-allelic homologous recombination in
repeat rich areas.

A small region of a chromosome is repeated, and placed next to the
existing copy (in tandem). duplications tend to happen in high
repetitive sequences of DNA. These repeated regions mean that the
chromosome will note line up very well with its homologous pair.

passed on by unequal order, and may or may not have a phenotypic effect.

\hypertarget{huntington-purea-1}{%
\paragraph{Huntington Purea}\label{huntington-purea-1}}

Repeats alter gene function, which can lead to brain damage. Repeats may
have a positive effect because they allow one copy of a gene to evolve
independently, specialise in a different function while the original
function is still conserved by the other copy of the gene.

\hypertarget{inversions.-1}{%
\paragraph{Inversions.}\label{inversions.-1}}

Cause: non-allelic homologous recombination Effects: 1. pairing in
meiosis occurs via looping 2. cross over in the looped region can cause
major deletions or duplications.

\hypertarget{translocation-1}{%
\paragraph{Translocation}\label{translocation-1}}

Effects: 1. novel chromosomes are generated 2. individuals are usually
infertile. 3. can result in aneuploidy (such as 14q21q inherited down
syndrome) 4.

\hypertarget{terminal-1}{%
\subparagraph{terminal}\label{terminal-1}}

A piece breaks off a chromosome and sticks to the end of a
non-homologous chromosome.

\hypertarget{reciprocal-1}{%
\subparagraph{reciprocal}\label{reciprocal-1}}

Exchange of pieces between non-homologous chromosomes \#\#\#\#
Robertsonian

two Acrocentric chromosomes lose their short arms and get stuck
together.

\hypertarget{fission-1}{%
\paragraph{Fission}\label{fission-1}}

deletion (loss of acentric piece)'

\hypertarget{fusion-1}{%
\paragraph{Fusion}\label{fusion-1}}

bicentric chromosomes (breakage in mitosis)

\hypertarget{breakage-fusion-bridge-cycles-in-cancer.-1}{%
\subparagraph{breakage fusion bridge cycles in
cancer.}\label{breakage-fusion-bridge-cycles-in-cancer.-1}}

Copy image. 1. end of a chromosome breaks off (during meiosis) 2. one
daughter cell inherits the chromosome lacking a telomere. 3. the new
cell enters S phase and replicates its DNA (it now has two sticky DNA
ends right next to each other. ) 4. loose DNA strands bind to form
continuous DNA loop. 5. fused sister chromatids are pulled apart during
mitosis, causing breakage at a new sight 6. one daughter cell inherits
chromosome with duplicated genes and again a free sticky end. this cycle
can then begin again leading to even further chromosome distortion.

\hypertarget{imprinting-1}{%
\paragraph{Imprinting}\label{imprinting-1}}

\hypertarget{hinny-vs-mule-1}{%
\subparagraph{Hinny vs mule}\label{hinny-vs-mule-1}}

Mule

horse mother donkey father hardy and obedient

Hinny

donkey mother, horse father. temperamental, untrainable.

\hypertarget{intersistial-deletion-of-chromosome-15-1}{%
\subparagraph{Intersistial deletion of chromosome
15}\label{intersistial-deletion-of-chromosome-15-1}}

Prader -Willi syndrome

Cause: deletion from father, ZNF127 and IPW inactivated on the maternal
copy.\\
effects: 1. mental retardation 2. obesity 3. diminished growth.

Angelman syndrome

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  deletion from mother: UBE3A inactivated on paternal copy.\\
\item
  mental retardation
\item
  epilepsy
\item
  lack of motor development.
\end{enumerate}

\hypertarget{x-inactivation-1}{%
\paragraph{X inactivation}\label{x-inactivation-1}}

the inactivation of one of two X chromosomes in females via DNA
methylation.

NOTE: which X chromosome is inactivated varies randomly from cell to
cell.this inactivation process turns the inactive X into a bar body,
(highly methylated and rolled up in histone proteins).

\end{document}
