---
title: "Discharge Calculation"
output: pdf_document
---

# Notes
assume endges are triangular. (w1*d1*1/2*v)

# Rondegat River

```{r, echo=FALSE}
# Rondegat River discharge calculation


# Variable declaration.

Dis<-c((1:8)/2)
Dep<-c(0.56,0.61,0.58,0.51,0.44,0.24,0.1)
V<-c(0.1,0.48,1.31,1.17,0.52,0.13,0.32)
V[8]<-c(V[7])

# Calculations. 
x<-(1:7)
W<-c(Dis[x+1]-Dis[x])

A<-c(Dep[x]*W[x])
A[1]<-c(A[1]/2)
A[8]<-c(Dis[8]*Dep[7]/2)


y<-(1:8)
sD<-c(A[y]*V[y])
fD<-sum(sD)
```

Dishcarge of each segment is \newline 

 `r {sD[1]}`$m^3\cdot s^{-1}$
 
  `r {sD[2]}`$m^3\cdot s^{-1}$
 
  `r {sD[3]}`$m^3\cdot s^{-1}$
 
 `r {sD[4]}`$m^3\cdot s^{-1}$
 
  `r {sD[5]}`$m^3\cdot s^{-1}$
 
  `r {sD[6]}`$m^3\cdot s^{-1}$

  `r {sD[7]}`$m^3\cdot s^{-1}$
 
  `r {sD[8]}`$m^3\cdot s^{-1}$
`

and the overall discharge is: \newline 

`r {fD}`$m^3\cdot s^{-1}$

# Magalies River 

```{r, echo=FALSE}

# Magalies River discharge calculation.

# Variable declaration.
Dis<-c(1:6)
Dep<-c(0.35,0.41,0.38,0.33,0.09)
V<-c(0,0.05,0.03,0,0)
V[6]<-c(V[5])

# Calculations
x<-c(1:5)
W<-c(Dis[x+1]-Dis[x])

A<-c(Dep[x]*W[x])
A[1]<-c(A[1]/2)
A[6]<-c(Dis[6]*Dep[5]/2)
length(A)

y<-(1:6)
sD<-c(A[y]*V[y])

fD<-sum(sD)
```


Dishcarge of each segment is \newline 

 `r {sD[1]}`$m^3\cdot s^{-1}$
 
  `r {sD[2]}`$m^3\cdot s^{-1}$
 
  `r {sD[3]}`$m^3\cdot s^{-1}$
 
 `r {sD[4]}`$m^3\cdot s^{-1}$
 
  `r {sD[5]}`$m^3\cdot s^{-1}$
 
  `r {sD[6]}`$m^3\cdot s^{-1}$


and the overall discharge is: \newline 

`r {fD}`$m^3\cdot s^{-1}$
