---
title: "Inverstigation into the deleterious impact of urbanisation on the freshwater ecosystem of the Braamfontein Spruit"

author: "Keelan Krinsky: _1634953_"
output: pdf_document
---

\footnote{27th October, 2018}

\newpage

# \textbf{ABSTRACT}
# \testbf{KEYWORDS}
Freshwater Ecosystems; Habitat Integrity; South Africa Scoring System; Urbanisation

# \textbf{INTRODUCTION}

Globally, urban freshwater ecosystems play a variety of very important social, economic and ecological functions. These systems function as: recreational areas (Swanwick _et al_, 2003); animal and plant habitats, an important source of fish stocks (Berkes, 1979), flood control systems (Levy _et al_, 2007), domestic or industrial water treatment and disposal systems; and even as a source of bathing or even drinking water in some communities. However, urban, freshwater systems come under considerable stress due to human disturbance, predominately in the form of pollution, so much so that there their natural functions may become entirely compromised. A detailed understanding of the nature and consequences of anthropic disturbance, may prove critical for effective protection and remediation of these vital systems.   

# NB: rewrite this in terms of: it was found that this pollutant with this human cuase had this effect. 

Much research has already been done into the deleterious effects of pollution on aquatic ecosystems. Nitrate pollution, deriving from domestic sewerage, has been linked, at a global level, to eutrophication, and the resulting losses in biodiversity and system function, (Julio and Alvaro, 2006). Other nutrient pollution, such as phosphate pollution from agricultural fertalisers, or washing powders has also been shown to decreased species diversity, nutrient uptake, and productivity within aquatic systems, (Walsh _et al_, 2005). Finally industrial pollution such as, acid mine drainage, can lead to increased heavy metal toxicity, which ???????? species are particularly sensitive to ( _Citation needed_), and can cause toxicity in drinking water ( _Citation needed_) (Ata and Soner, 2006).

Physical sources of disturbance have also been assessed, such as the effects of storm water run off, in which a connection was established  between water run off schemes and river ecological condition. (Walsh _et al_, 2012). Erosion from ground disturbance in biulding can also wash down into urban aquatic systems increasing the turbidity, and potentially even blocking up the water channel completely ( _Citation needed_). 

Biological source of disturbance, may also be anthropically induced and can act as very serious disrupt a system. Introduction of invasives such as ????? ( _citations needed_), have again lead to decreases in biodiversity and water quality. 


 Overall, many pertinant anthropic causes of pollution have been identified and categorised, and their effects inverstigated; however, due to the underlying biological complexity of the issue new polutants, and new biological consequences of previously identified pollutants, are still continuously being identified. ( _citation needed_). Although a strong background has been developed for assessing the damage related to urbanisation and anthropic pollution, much additional research will still be required to fully detail the extent and complexity of the issue. 

The aim of this study was to assess the effects of urbanisation and anthropic disturbance on the ecological state on of the Braamfontein Spruit freshwater ecosystem. This aim was split into several objectives. Firstly, to assess the extent, and nature of physiochemical changes within the urban  system, specifially, dissolved oxygen, pH, Temperature and turbidity. Secondly to assess the differences in biological diversity, and by extention water quality in the urban versus the rural system, using the South African Scoring System Version 5 (SASS) methodology. Thirdly to assess any differces of the ecological integrity of the surrounding catchment between the urban and the rural system using Index of Habitat Integrity (IHI) scoring methodology. The SASS metholodology uses the diversity of macroinvertibrate species observed as an indicator of water quality, as many rarer macroinvertibrate species are dependant on high water quality for survival. The IHI scoring methodology relies on visual clues of anthropic presence within the catchement to access the overall level of anthropic disturbance.  


The aim of this investigation was to establish the extent, and origin, of human generated pollution and disturbance on the Delta Park Braamfontein spruit, to assist in the mapping of freshwater pollution across the Johannesburg city area. Within this general aim objectives included Establishing the major sources of disturbance and their origin, as well as assessing the overall ecosystem health as measured by the SASS scoring methodology, and finally to assess the magnitude of the disturbance within the catchment by comparison to the prestine Magalies headwaters system, located in the same general geographic area but suffering from far less human generated pollution.  
\newpage

# \textbf{METHODS AND MATERIALS}

## \textbf{Study area}
Two separate sites were assessed in this study, a lowlands section of the Braamfontein Spruit (the main study site) lying within a highly urbanised area, and a prestine upland section of the Magalies River, used as a refereance site. 


## \textbf{Braamfontein Spruit study site}
The Braamfontein Spriut study site shown in Figure 3 lies within, Delta Park, a public park located along Craighall Road within Victory Park, Johannesburg, a city within the Provence of central gauteng. 

The site is prone to frequen serious erosion due to flooding from storm water drains. (Masawi, 2018)

GPS coordinates $26^\circ07'35.3"S 28^\circ00'59.1"E$
The parks lies within a highly populated urban district with many recreational users frequenting the park every day, as well as homeless individuals who inhabit the park, and may use the river waters for bathing or cooking. Furthermore the park was historically a sewerage treatment site any although it has been re-purposed, it is still a hub of the Johannesburg sewerage transport network the the possibility of multiple leaks of raw sewerage directly into the river, as well as storm water drains which empty directly into the river. For assessment of water quality the SASS scoring based on the diversity of aquatic macro-invertebrates invertebrate found, was assign to each sample site (Dickens and Graham, 2002). A relatively pristine section of the Magalies river was also sampled as a control/stand for comparison.




The Magalies River study site, shown in Figure 3, lies just of the ?? Highway within the ??? region in the Magalies Mountains just outside of Johanessbur City. 

## \textbf{Experimental design and protocol 

### \textbf{Physiochemical parameters}
For each study site, measures of physiochemical parameters and biological/habitat indicators were assessed at four sample sites displaced $\approx$ 30m from eachother along the stream channel. For each sample site an indentical sampling porceedure was employed. 

Before any other sampling took place, a turbidity sample was taken by gently submerging a sampling bottle in the center of the channel, and allowing it to fill. The sampling bottle was then sealed tightly, for storage and taken for laboritory analysis. A set of dilutions were made of the sample and the absorbance of each dilution measured in Nephelometric Turbidity Units using a Hanna turbidity meter (model HI98703-02). 

pH, Temperature, dissolved oxygen, and conductivity were all assessed using a handheld YSI Professional Plus multi-parameter meter; measurements were taken by placing the probe in the center of the water channel, and taking the measurment only once the reading stabalised. Only one reading was taken at each sample site.   

A Transparent Velocity Head Rod (TVHR) (sourced from the company GroundTruth, operating in Pietermaritzburg, South Africa), along with a 50m tape measure were used to make the measurements required to calculated discharge. The tape measure was used to measure the in channel width of a slow flowing section of the channel with a smooth, homogeous benthos, i.e an area with no sudden variations in high or large stones on the river bed, or macrophytes blocking the river flow.  The TVHR was then used to, measure the depth of the channel at 0.5m intervals along the horizontal line which was used in measuring channel width. The ??? flow,??? and were also measured and used to calculate an estimate of water velocity at this point in the channel. 

### \textbf{Biological indices}
The Standard Methodolgy described in the SASS version 5 was used for the Sampling of Aquatic invertibrates with some small changes. Instead of (??? nets), butterfly catching nets were used, and instead of the 2 minute sampling period indicated, in cases where it appeared few vertibrates had been collected the period was extended to five minutes. 

The invertibrate samples, once collected where stored in $\approx$ 85\% ethanol solution, and stored for several weeks, before being identified to family level. 

### \textbf{Habitat integrity}
Signs of human impact were observed and categorised in nature and extent by using the ???? system as in ???? reference. 


## \textbf{Data Analysis}
The overall means of each physiochemical measure, as well as the habitat integrity score, and biological indicators were calculated from the data collected at each sample site, to reduce no-determinate errors, and local variations associated indivdual measures. The standard error for each parameter was also calculated to given an indication of the natural variation seen, and the precision of the measurements obtained. 

Two sample, equal varience, two tailed, Student t tests were also performed to test for significant differences in each parameter between the two study site (Braamfotein spriut and Magalies River), at a 5\% signifcance level. Equal varience of the two study site was assumed with no explicit statistical test. Two tailed tests were used as the purpose of the study was simply to assess the difference between disturbed and prestine habitats, not to test any specific theory of the effects of disturbance on biological and physiochemical conditions. 

# \textbf{RESULTS} 

## \textbf{Physiochemical data}
Summary measures of the physiochemical data collected are shown in Table 1, while statistical analysis of the physiochemical differences between the two sampling sites is shown in Table 2, below. 

In Table 1 it can be seen that measures of Conductivity, Dissolved oxygen, and Turbidity were all a good deal higher, on average at the Braamfontein Spriut sampling sight, at which there was a higher current, and more riffles/shallow rocky sections, and more macrophytes growing within the water channel itself. Manholes for operational sewers were also located near to the banks of the stream at this site. Furthermore, it can be seen from channel two that the differences observed in Turbidity, dissolved oxygen concentration were all both highly significant well below the conventional 5\% significance level set for the purpose of this study. Conversely, however, the difference observed in average conductivity had only a low associated significance level (Predominately due to the very large variance in the conductivity measures taken), and so was not considered sifnificant for the purposes of this study. 

It can also be seen from Table 1 that measures of pH and Temperature, were higher at the Magalies sampling site, however only the difference in temperature actually proved to be significant, as seen in Table 2.

## \textbf{Habitat assessment}
As seen in Table 1 the Index of Habitat Integrity (IHI) score was far higher at the Braamfontein spruit site, which as afor mentioned, in a densly populated urban area, and exhibited visible signs of human disturbance such as plastic waste floating within the channel itself, as well as large storm drain pipes draining dirrectly into the channel and bridges and a large pedestrian bridge placed over the channel. This significant difference is futher demonstatred by Figure 1, in which it can be seen that not only is the mean IHI score at the Braamfontein spriut sampling site more than double that of IHI score for the Magalies River sampling site, there is is no overlap in the error bars for the two means, Theses error bars, which show the standard error in the means calculated support the observation of significant difference given in table 2, showing that the difference in means is highly unlikely to have assisen simply from random variation.  

## \texbf{Biological indicators}
All of the indicators of bioligical diversity measured i.e, South Africa Scoring System, version 5 (SASS) Score, taxa score, and Average Score per Taxon (ASPT), (Also calculated with the SASS method), were higher at the Magalies sampling site at which different biome types were more distinctly segregated, and evenly distributed along the river channel, as seen in Table 1. The differences observed in these parameters between the two sampling sites were without exception significant, as seen in Table 2.  

```{r, echo=FALSE}
# Data Input
se <- function(x) sqrt(var(x)/length(x))


## Brammfontein Spruit
bpH<-c(8.1,8.26,8.12,7.94,8.34,8.39,8.31)
bC<-c(366.3,367.1,150.9,151.5,343.9,356,148.7)
bTe<-c(15,15,13.9,13.8,13.9,15.3,13.9)
bDO<-c(8.22,"NA",8.82,"NA",10.63,8.46,9.33)
bDOe<-c(8.22,8.82,10.63,8.46,9.33)
bTu<-c(13.1,6.23,12.1,20.5,18.3,13.7,11)
bAv1<-c(mean(bpH), mean(bC), mean(bTe), mean(bDOe), mean(bTu))
bse1<-c(se(bpH), se(bC),se(bTe), se(bDOe), se(bTu))


bIHI<-c(66,122,123,70," ",175,163,167)
bIHIe<-c(66,122,123,70,175,163,167)
bSASS<-c(55,53,38,30,47,32,48,51)
bt<-c(12,11,8,8,9,7,11,9)
bASPT<-c(4.58,4.82,4.75,3.75,5.2,4.57,4.36,5.7)
bAv2<-c(mean(bIHIe), mean(bSASS), mean(bt), mean(bASPT))
bse2<-c(se(bIHIe),se(bSASS), se(bt), se(bASPT))
bsd<-sd(bIHIe)


Bp<-data.frame("pH"=bpH, "Conductivity"=bC, "Temperature"=bTe, "Dissolved Oxygen"=bDO,"Turbidity"=bTu)

Bb<-data.frame("IHI"=bIHI, "SASS"=bSASS, "taxa"=bt, "ASPT"=bASPT)

bAv<-c(bAv1,bAv2)
bse<-c(bse1,bse2)

## Magalies

mpH<-c(8.36,	8.31,	8.23,	8.31,	8.42,	8.34,	8.25)
mC<-c(224.8,	220.8,	226.7,	224.8	,226.1,	222.7,	225.6)
mTe<-c(19,	18.4,	19,	18.9,	19.2,	18.6,	19)
mDO<-c(5.76,	6.52,	6.54,	6.54,	6.55,	6.09,	5.5)
mTu<-c(1.33,	0.83,	1.24,	2.04,	0.98,	2.13,	2.65)
mAv1<-c(mean(mpH), mean(mC), mean(mTe), mean(mDO), mean(mTu))
mse1<-c(se(mpH),se(mC),se(mTe),se(mDO),se(mTu))



mIHI<-c(7,	42,	27,	162,41,	2,	15)
mSASS<-c(216,	193,	80,	108,	139,	78,	100,	61)
mtaxa<-c(31,	26,	13,	16,	24,	14,	17,	10)
mASPT<-c(6.97,	7.42,	6.15,	6.75,	5.79,	5.57,	5.88,	6.1)
mAv2<-c(mean(mIHI), mean(mSASS), mean(mtaxa), mean(mASPT))
mse2<-c(se(mIHI),se(mSASS), se(mtaxa), se(mASPT))
msd<-sd(mIHI)

mAv<-c(mAv1,mAv2)
mse<-c(mse1,mse2)

mAv<-round(mAv,digits=2)
mse<-round(mse,digits=2)

bAv<-round(bAv,digits=2)
bse<-round(bse,digits=2)


```

```{r, echo=FALSE}
# Table of water quality data 

x<-c(1:9)
mse[x]<-c(paste("(",mse[x],")"))


bse[x]<-c(paste("(",bse[x],")"))


T1<-data.frame("Mean"=bAv,"Standard Error"=bse,"Mean "=mAv,"Standard Error "=mse)
library(knitr)


row.names(T1)<-c("pH","Conductitiy ($\\mu S$)","Temperature($^\\circ C$)","Dissolved Oxygen($mg \\cdot L^{-1}$)", "Turbidity(NTUs)","IHI", "SASS", "taxa", "ASPT")

library("kableExtra")
T1d_latex<-knitr::kable(T1,"latex", digits = 3,row.names=NA, col.names = NA, align="llll", caption = "Summary measure of key physiochemical characteristics, biological indicators of water quality, and habitat integrity indicators for the Braamfotein spruit study site, and Magalies river prestine reference site.", escape = FALSE) %>%
add_header_above(c("", "Braamfontein Spruit"=2, "Magalies River"=2))


kable_styling(T1d_latex, latex_options = c("striped","hold_position"), position= "center")


```

```{r, echo=FALSE}
### Comparison table. 

pv<-c(t.test(bpH,mpH,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bC,mC,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bTe,mTe,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bDOe,mDO,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bTu,mTu,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bIHIe,mIHI,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bSASS,mSASS,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bt,mtaxa,alternative=c("two.sided"), var.equal = TRUE)$p.value,
     t.test(bASPT,mASPT,alternative=c("two.sided"), var.equal = TRUE)$p.value)
     
df<-c(t.test(bpH,mpH,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bC,mC,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bTe,mTe,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bDOe,mDO,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bTu,mTu,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bIHIe,mIHI,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bSASS,mSASS,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bt,mtaxa,alternative=c("two.sided"), var.equal = TRUE)$parameter,
     t.test(bASPT,mASPT,alternative=c("two.sided"),var.equal = TRUE)$parameter)


x<-data.frame("x"=df)
df<-c(x$x)


ts<-c(t.test(bpH,mpH,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bC,mC,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bTe,mTe,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bDOe,mDO,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bTu,mTu,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bIHIe,mIHI,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bSASS,mSASS,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bt,mtaxa,alternative=c("two.sided"), var.equal = TRUE)$statistic,
     t.test(bASPT,mASPT,alternative=c("two.sided"), var.equal = TRUE)$statistic)




x<-data.frame("x"=ts)
ts<-c(x$x)
```

```{r, echo=FALSE, warning=FALSE}
## Table 2 statisitcal comparison. 

pv<-round(pv,digits=2)

T2<-data.frame("T statistic"=ts,"Degrees of freedom"=df, "p value"=pv )
library(knitr)


row.names(T2)<-c("pH","Conductitiy ($\\mu S$)","Temperature($^\\circ C$)","Dissolved Oxygen($mg \\cdot L^{-1}$)", "Turbidity(NTUs)","IHI","SASS","taxa","ASPT")


library("kableExtra")
T2d_latex<-knitr::kable(T2,"latex", digits = 3,row.names=NA, col.names = NA, align="llll", caption = "Student t test, for significant difference in physiochemical characteristics, biological indicators of water condition, and measures of habitat integrity between the Braamfontein Spruit sample site, and the prestine Magalies river referecen site.", escape = FALSE) 
kable_styling(T2d_latex, latex_options = c("striped","hold_position"), position= "center")

```

\newpage

## \textbf{Biological indicies}
Figure 1 below shows the comparison between the IHIH score calculated for each site, and figure 2 illustrates the relation between (average) ASPT and SASS, to classically the overall state of the system. 

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
bars<-c(bAv2[1],mAv2[1])


bsse<-c(bsd[1]/((length(bIHIe))^(1/2)))
msse<-c(msd[1]/((length(mIHI))^(1/2)))



err<-c(bsse[1],msse[1])



df<-data.frame("bars"=bars,"err"=err)
row.names(df)<-c("Braamfontein Spruit","Magalies River")

ggplot(df, aes(x=row.names(df), y=bars), xlab="Sample Site") +
  geom_bar(stat="identity", fill="grey") +
  geom_errorbar(aes(ymin = bars-err, ymax = bars+err),width=0.1) +
  theme_classic()+
   xlab("Sampling Sites") +
  ylab("IHI Score") +
    ggtitle(expression(" \n \n \n Figure 1: Index of Habitat Integrity (IHI) scores for the Braamfotein spruit study site catchment, and Magalies river prestine reference site catchment.")) +
     theme(plot.margin = unit(c(1,1,3,1), "cm"))+
  theme(plot.title = element_text(hjust=.6,vjust=-90, size=11))
```

$\quad$\newline$\quad$\newline
$\quad$\newline$\quad$\newline


```{r, echo=FALSE, warning=FALSE}
## Scatter plot Braam
AvA<-c(bAv2[4], mAv2[4])
sd1<-c(sd(bASPT), sd(mASPT))

AvA1<-c(bAv2[2], mAv2[2])

AvS<-c(bAv2[2], mAv2[2])
sd2<-c(sd(bSASS), sd(mSASS))


AvS2<-c(AvS,mAv2[2]-mse2[2], bAv2[2]-bse2[2])
AvS1<-c(AvS,mAv2[2]+mse2[2], bAv2[2]+bse2[2])


lAvS<-c(AvA-sd1,mAv2[4],bAv2[4])
uAvS<-c(AvA+sd1,mAv2[4], bAv2[4])

par(mar = c(10,10,2,2) )
plot(AvS,AvA, xlab="SASS",ylab="Average Score per Taxon", ylim=range(c(3,8)), xlim=range(c(30,140)))
arrows(AvS2, lAvS, AvS1, uAvS, length=0.05, angle=90, code=3)
arrows(88,0,88,4.5,length=0,col="blue")
arrows(115,0,115,5.4,length=0,col="blue")
arrows(151,0,151,5.9,length=0,col="blue")
arrows(244,0,244,6.4,length=0,col="blue")

arrows(0,4.5,88,4.5, length=0,col="blue")
arrows(0,5.4,115,5.4, length=0,col="blue")
arrows(0,5.9,151,5.9, length=0,col="blue")
arrows(0,6.4,244,6.4, length=0,col="blue")


text(70,3.7,"E/F")
text(100,4.8,"D")
text(130,5.6,"C")
text(200,6.2,"B")
title(sub="Figure 2: The average score per taxon (ASPT) vs the overall Ecosystem \n score, calculated using the South African Scoring system \n for rapid bioassessment (SASS-5), Version 5.\n The zones donated by letters E to B indicate categories of ecosystem integrity from A the least degraded, to most E the most degraded",cex.sub=1, line=7 )
```


# \textbf{DISCUSSION} 

The effects of urbanisation are clearly present across all aspects of the Braamfontein spruit system, from the physiochemical characteristics, to the biological indicators, to measures of the integrity of the surrounding habitat. Furthermore it is more than likely that changes are closely connected, and linked to the degradation of the ecosystem as a whole. Arguably the process underlying the degradation observed may originate with changes in the physiochemical characteristics related to anthropic disturbace. 

The significant difference in temperature observed was unexpected as mountanous headwater are conventionally expected to be cooler than larger lowland streams. ( _Citation needed_). The lower temperature in the Braamfontein spruit may related to the increased turbidity, as even at very shallow depth below the surface the penetration of sunlight, and the resultant heating can be markedly decreased ( _citation needed_). However given the vigorous current, the exposure to wind action (due to low tree cover) and the frequent riffles present in the Braamfontein spruit physical/mechanical water mixing might be expected bring the sunlight heated surface waters into thermal equilbirum with the water just below minimizing the effects of this differece. Perhaps a more likely explanation is that in influx of cold storm water, bringing in water which enters the river dirrectly from the underground storm sewers system (Hence not exposed to sunlight), may have resulted in thermal pollution lowering temperature of the Braamfontein Spruit site ( _Citation for thermal pollution_).

The significant difference in oxygen concentration observed may well be a dirrect result of the temperature difference seen, as oxygen solubility in water increases markedly with decreased temperature. ( _Citation Needed_). 

The decreased biological diversity indicated by significant changes in SASS-5 score is a biological indicator of poor water quality. This poor water quality may be due, at least in part to the changes in physiochemical condition seen. Increased turbidity, for instance, can have detrimental effects of many macro invertibrates such as ????????????? ( _Citation_). Temperature variation related to heat pollution may also have an impact on biodiversity, with large variations in temperature leading to decreased abundance of stenotherm species such as ??????. 



Decreased biodiversity and ecological condition may also to degradation of the surrounding catchment. 

Catchmet degradation may also play a role in the physiochemical differences seen. The increase in tubidity may well be related to the increased area of impervious surfaces (i.e) tar rodes and paving within the catchment preventing perculating and leading to run off as a pose to subsurface flow, which is more likely to pick up loose sediment and deposit it into the river channel. Fine building sand dumped on the pavements, of surrounding houses, or or mud and clay from excavations in road works will also be picked up by run off and accumulate within the river channel adding to the sediment load. 


# \textbf{ACKNOWLEDGEMENTS}
University of the WitsWaterstrand for providing all sampling material, as well as reference books used in macroinvertibrate identification for SASS scoring. 

The Witswaterstand Students who aided in all sampling and data collection at both study sites. 


# \textbf{REFERENCES}

Ata A., Soner K.,2006, Acid Mine Drainage (AMD): causes, treatment and case studies,_Journal of Cleaner Production_,14,12–13: 1139-1145,

Berkes, F., 1979, An Investigation of Cree Indian Domestic Fisheries in Northern Quebec. Arctic, 32(1), 46-70.

Christopher J. Walsh, Allison H. Roy, Jack W. Feminella, Peter D. Cottingham, Peter M. Groffman, and Raymond P. Morgan,2005, "The urban stream syndrome: current knowledge and the search for a cure," _Journal of the North American Benthological Society_ 24,3: 706-723.

Dickens C.W.S., Graham P.M., 2002, The South African Scoring System (SASS) Version 5 Rapid bioassesment method for rivers,Umgeni Water, _African Journal of Aquatic Science_ ,27:1-10

Julio A. Camargo, Álvaro Alonso,2006, Ecological and toxicological effects of inorganic nitrogen pollution in aquatic ecosystems: A global assessment, _Environment International_, 32,6,2006: 831-849,

Levy, J. K., Hartmann, J. , Li, K. W., An, Y. and Asgary, A. (2007), Multi‐Criteria Decision Support Systems for Flood Hazard Mitigation and Emergency Response in Urban Watersheds. _ Journal of the American Water Resources Association_, 43: 346-358. 

SWANWICK, C., DUNNETT, N., & WOOLLEY, H. 2003. Nature, Role and Value of Green Space in Towns and Cities: An Overview. Built Environment (1978-), 29(2), 94-106.

Sabo, J. L., Sponseller, R. , Dixon, M. , Gade, K. , Harms, T. , Heffernan, J. , Jani, A. , Katz, G. , Soykan, C. , Watts, J. and Welter, J. (2005), RIPARIAN ZONES INCREASE REGIONAL SPECIES RICHNESS BY HARBORING DIFFERENT, NOT MORE, SPECIES. _Ecology_, 86: 56-62.

Walsh CJ, Fletcher TD, Burns MJ, 2012, Urban Stormwater Runoff: A New Class of Environmental Flow Problem. _PLOS ONE_ 7(9)


Ndompupei Masawi, Report number, 490343, prepared for Johannesburg Roads Agency (JRA), srk consulting, 2018, Johannesburg South Africa. 

\newpage

# Questions

Only mention SASS as why its important in introduction i.e how its useful. 

what could be improved

looked at riverine habitat, underlying geology. 

the question of temperature

study site materials all in the study site. 

IHI gives condition of catchment of the area. In the land which feeds the river. percolate and subsurface flow tends to be a lot cleaner. 

physiochemical in introd think of reader. 

 #. Must we have GPS coords for all four sites. 
 #. should we show map and image. 
 #. what are the model numbers of the pH probe and thermometer. 
 #. should data be qouted in the discription section of the results. 
 #. was conductivity measured or estimated. 
 #. must i reference SASS twice in the methods and in the introduction.
 #. where do observations of alien plants and ducks belong, results or site discription. 
 #. Are we testing an hypothesis here, the t tests seem to indicate that we are but surely this is not an issue that is debated (i.e the fact that urbanisation causes degradation.)


# Posible imporvements 
 #. break up and expand the introduction, i.e there can be a paragraph on eutrophication, one on metal toxicity etc. 