---
title: "Tut Problems"
output: html_notebook
---

## Tutorial one. 

### Question 1. 

column 1  Spectral region
 #. Visible
 #. microwaves
 #. radiowaves.
 #. Infrared 
 #. Ultra Violet. 

 column 2 Type of transition. 
 
 #. Electron
 #. 
 #. Electron of proton spin magnetic transitions. 
 #. Bond stretch and bending
 
 Column 3 $\lambda$
 
 #. 500nm
 #. 3.00cm
 #. 3m
 #. 0.01mm
 #. 239nm
 
 Column 4 $v/Hz$
 
 #. $6.00 \cdot 10^{14}$
 #. $1\cdot 10^{10}$
 #. $10^8$
 #. $3\cdot10^{13}$
 #. $1.25 \cdot 10^{15}$
 
 
 Column 5 $\bar{\nu}/cm^{_-1}$
 
 #. $2.00 \cdot 10^{4}$
 #. $3.33 \cdot 10^{-1}$
 #. $3.33 \cdot 10^{-3}$
 #. $1 \cdot 10^3$
 #. $4.18 \cdot 10^4$
 
 Column 6 $E/J$
 
 #. $3.97 \cdot 10^2$
 #. $6.626 \cdot 10^{-24}$
 #. $6.626 \cdot 10^{-26}$
 #. $1.988 \cdot 10^{-20}$
 #. $8.30 \cdot 10^{-19}$
 
 Column 7 E/ kJ \cdot mol^{-1}
 
 #. $2.39 \cdot 10^2$
 #. $3.99 \cdot 10^{-3}$
 #. $3.99 \cdot 10^{-5}$
 #. $11.97 \cdot 10^{1}$
 #. $5.00 \cdot 10^2$
 
### Question #2

```{r, ehco=False}
e<-c(3*10^8/(300*10^(-9)))*6.626*10^(-34)
e

b<-c(299/(6.022* 10^(20)))
     
```
Energy of photon= `r {e}`
Enegry of bond =`r {b}`

Yes as a single photon of the incident light contains more than enough energy to split the H-I bond if it were absorbed. Accurate results cannot be obtained becuase during the measurment process HI will be converted to $H^{+\cdot}$ and $I^{+\cdot}$  which will have there own absorption spectra which will affect the overall absorption of the sample. 

### Questions #3
$A\-log(T)$

```{r}
T1<-round ((10^(-0.1)), digits=1)
T2<-round ((10^-(0.5)), digits=1)
T3<-round ((10^-(1)), digits=2)
T4<-round ((10^-(2)), digits=2)
T5<-round ((10^-(3)), digits=2)
```

`r {T1}`, `r {T2}`, `r {T3}`, `r {T4}`. `r {T5}`.

### Question #4 

#### A 
$C_4H_6$. 

$DBE=\frac{H_t-H_a-Hal+N}{2}=\frac{10-6}{2}=2$

Potential function groups \newline
cycloalkene, or alkene.  

#### B 
$C_{10}H_{20}O_2$ 

$DBE=\frac{H_t-H_a-Hal+N}{2}=\frac{22-20}{2}=1$

Potential function groups \newline

 #. cycloalkane
 #. alkene 
 #. ketone 
 #. aldehyde
 #. carboxylic acid. 
 #. ester. 
(may contain an ether, or a diether, or a alcohol)

#### C 
$C_{6}H_{11}NO$ 

$DBE=\frac{H_t-H_a-Hal+N}{2}=\frac{14-11+1}{2}=2$

Potential function groups \newline
 #. alcohol
 #. amine
 #. amide
 #. cyloalkane
 #. nitrile
 #. nitroso 
 
#### D
$C_{8}H_{8}BrCl$ 

$DBE=\frac{H_t-H_a-Hal+N}{2}=\frac{18-8-2}{2}=4$

Potential function groups \newline

 #. benzene 
 #. bromoalkane (/ene)
 #. chloroalkane (/ene)
 #. cycloalkane (/ene)
 #. alkyl
 #. alkene
 
## Part 2 
 
### Question 3 A. 

#### Spectrum A
 #. peak one, $\approx 2900cm^{-1}$ C-H of alkane
 #. peak two $\approx 2800cm^{-1}$ C-H of aldehyde aldehyde
 #. peak three $\approx 2690cm^{-1}$ C-H stretch of aldehyde  
 #. peak two $\approx 1720cm^{-1}$ C=O stretch of an aldehyde
 
 Spectrum a corresponds to hepthanal. 
 
#### Spectrum B 
 
 #. peak one $\approx 3150cm^{-1}$ (broad) bonded O-H stretch in a promary alcohol   
 #. peak two $\approx 1710cm^{-1}$ C=O stretch in ketone. 
 #. peak three $\approx 1050cm^{-1}$ C-O stretch of a primary alcohol. 
 
 Spectrum B corresponds to acetic acid. 
 
#### Spectrum C 

 #. Peak one $\approx 3400cm^{-1}$ bonded O-H stretch of secondary alcohol
 #. Peak two $\approx 2970$ C-H stretch in alkane. 
 #. Peak three $\approx 2930cm^{}-1$ C-H stretch in alkane. 
 #. Peak four $\approx 2870cm^{-1}$ C-H stretch in alkane. 
 #. Peak five $\approx 1160cm^{-1}$ C-O stretch in secondary alcohol. 
 
 Spectrum C corresponds to 2-butanol. 
 
#### Spectrum D 
 #. peak one $\approx 2990cm^{-1}$ C-H stretch. 
 #. peak two $\approx 2940cm^{-1}$ C-H stretching in alkanes
 #. peak three $\approx 2900cm^{-1}$ C-H stretch in alkane
 #. $\approx 1705cm^{-1}$ C=O of a ketone. 
 
 Spectrum D corresponds to a 3 pentone. 
 
 
### Question 4

#### Spectrum 1 

 #. Peak one $\approx 3400$ bonded O-H stretch of an alcohol. 
 #. peak two $\approx 2950$ C-H stretch in alkane. 
 #. peak three $\approx 2810$ C-H of an aldehyde/ 
 #. peak four $\approx 2750cm^{-1}$ C-H stretch of an aldehyde. 
 #. peak five $\approx 1740cm^{-1}$ C=O of an aldehyde. 
  
The functional groups present are an alcohol and an aldehyde. 

#### Spectrum 2

 #. $\approx 3400cm^{-1}$ bonded O-H stretch of an alcohol.
 #. $\approx 2970cm^{-1}$ C-H stretch of an alkane. 
 #. peak three $\approx 2880cm^{-1}$ C-H stretch of an alkene. 
 #. peak four $\approx 1060cm^{-1}$ C-O stretch of a primary alcohol

The functional group present is a primary alcohol. 
 
#### Spectrum 3

 #. peak one $\approx 3300cm^-1$ C-H of an alkyne. 
 #. peak two $\approx 2990cm^{-1}$ C-H of a alkane. 
 #. peak three $\approx 2880cm^{-1}$ C-H stretch of an alkane. 
 #. peak four $\approx 2100cm^{-1}$ C$\equiv$C of an alkyne. 
 
 Therefore the functional group contained is an alkyne, (not ethyne).
 
 
