---
title: "Scientific Practise: Bacteriophage Replication"
author: "Keelan Krinsky: 1634953"
output:

  pdf_document: 

  template: /home/keelan/University-Notes.-/resources/Templates/fullpage.tex

---

# Burst Size of T4 phage in _Escherichia Coli_, as determined by the Ellis and Delbruck one step growth curve. 

# Aim
The aim of this experiment was to determine the burst size of T4 phage in _Escherichia Coli_ . Within this aim the ojectives included constructing a sufficiently dilute bacterial and phage solutions 

# Results 
The average number of Plaques relating to each time experimental time interval was calculated and plotted in __Figure:1__ below. The full set of results can be seen in __Table:1__ of the appendix.  
\newline \newline
```{r,echo=FALSE, warning=FALSE}
Av<-c(16,54,100,238,12)
t<-c(10,20,30,40,45)
Cr<-c(Av*100^2*10)
Crp<-c(Cr/10^6)

Lines <- list(bquote(' One Step Growth Curve of bacteriophage T4 in '*italic(Eschericia~Coli)),bquote("second line x"^2))


#library("ggplot2")
#G1<-data.frame(Av,t)
#ggplot(G1,aes(Av,t))

#qplot(Av,t,main="x",sub="x")
#plot(t,Av,sub=do.call(expression, Lines))
                               
                               
#bold(Figure:1)*' One Step Growth Curve of bacteriophage T4 in '*italic(Eschericia~Coli)* ' Bacterial Host'))

ylab<-bquote('Concentration of Phage ('~10^6~ 'Phage/ml)')
par(mfrow=c(1,1), mar=c(7,7,7,2))
plot(t,Crp,type="l",pch=21, xlab="Time after infection (min)", ylab=ylab,xlim=range(0:50))
grid(10,10, lty="solid")
#abline(v=(seq(8,45,0.2)), col="lightgray", lty="solid",lwd=0.000001)
#abline(h=(seq(0,25,0.2)), col="lightgray", lty="solid", lwd=0.000001)
points(t, Crp, type="p", pch=1, col="black")
labels.default("Shit",5,5)




exp <- 2
Lines <- list(bquote('in '*italic(Eschericia~Coli)*' Bacterial Host'), bquote(bold(Figure:1)*' One Step Growth Curve of bacteriophage T4'))
mtext(do.call(expression, Lines),side=1,line=6:5 )
#lines(predict(lo), col='red', lwd=2)

#plot(1:10, main=(paste(expression('C=1'*alpha*'d'))))

```





## Burst size. 


# Discussion.

# Appendix1: Results

```{r, echo=FALSE}
S1<-c(21,6,11,26,23,25,11,10,14,16)
S2<-c(67,54,53,40,68,55,34,64,51,54)
S3<-c(131,92,128,87,93,99,92,93,89,100)
S4<-c(181,504,182,121,130,304,316,202,207,238) 
S5<-c(10,0,11,0,38,0,26,"NA","NA", 12)
RN<-c("Experiment1","Experiment2","Experiment3","Experiment4","Experiment5","Experiment6","Experiment7","Experiment8","Experiment9","Average" )

T1<-data.frame("RN"=RN,"1"=S1, "2"=S2, "3"=S3, "4"=S4, "5"=S5)

library(knitr)
colnames(T1)<-c("Time From start", "10 min","20 min", "30 min","40 min","45 min")



library("kableExtra")

T1d_latex<-knitr::kable(T1,"latex",row.names=NA, col.names = NA, align="ccccc", caption ="Record of Plaques obersved after $E. coli$ culitivation which, phage solution generated by forced lysis after the given time intervals ", escape = FALSE)

kable_styling(T1d_latex, latex_options = c("striped","hold_position", "bordered", position="center"))

```


# Appendix 2: Calculations
As each plaque forming unit is assumed to be a single phage the number of phage present after dilution can be taken as equal to the number of plaques counted. 

The concentration of phage in tube II is thererfore equalvent to $\frac{n}{0.1ml}$, where n is the number of phage counted. similiarty
concentration in tube =$\frac{n\,\, \text{x}\,\, 100}{0.1ml}$,and cencetration in ADS tube 2 =$\frac{n\,\, \text{x}\,\, 100^2}{0.1ml}$ which implies that the number of phage present in the reaction mixture after the given time intervals are as follows. 



```{r, echo=FALSE}
Cr<-c(Av*100^2*10)
S1<-c(Cr[1])
S2<-c(Cr[2])
S3<-c("10000000")
S4<-c(Cr[4])
S5<-c(Cr[5])
RN<-c("Average Concentration (Phage/ml)")



T2<-data.frame("RN"=RN,"1"=S1, "2"=S2, "3"=S3, "4"=S4, "5"=S5)

library(knitr)
colnames(T2)<-c("Time from infection", "10 min","20 min", "30 min","40 min","45 min")

library("kableExtra")

T2d_latex<-knitr::kable(T2,"latex",row.names=NA, col.names = NA, align="ccccc", caption ="Concentration of phage in original infection solution", escape = FALSE)

kable_styling(T2d_latex, latex_options = c("striped","hold_position", "bordered", position="center"))

```


The Initial concentration of phage in the reaction mixture= $\frac{2\cdot 10^8 \text{Phage} \cdot ml^{-1}}{100^2}=2.0\cdot 10^4$