\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother


\date{}

\begin{document}

\hypertarget{zeolites-i3}{%
\subsection{Zeolites I3}\label{zeolites-i3}}

\hypertarget{definitions}{%
\subsubsection{Definitions}\label{definitions}}

\hypertarget{zeolites}{%
\paragraph{Zeolites}\label{zeolites}}

crystalline aluminosiliates that are built up of tetrahedral
arrangements of \(SiO_4\) and \(AlO_4\) units

\hypertarget{construction}{%
\subsubsection{Construction}\label{construction}}

\(Al(OH_4)^-\) and \(Si(OH_4)\) units . the negative Aluminium oxide is
associated with a positive ion. if this ion is \(H^{+}\) then the
material is an acid, if it is a group 1 or 2 metal then these ions may
be exchanged with other ions. (Si and Al units can be condensed to form
dimers)

NOTE: regular arrangement at angstrom level. 1. molecular sieve. 2. acid
catalyst. 3. ion exchange (with the cations associate with the
aluminium)

\hypertarget{procedure}{%
\section{procedure}\label{procedure}}

\hypertarget{preparation}{%
\subsection{preparation}\label{preparation}}

\hypertarget{task-one}{%
\subsection{task one}\label{task-one}}

prepare 0.01 \$ CaCl\_\{2\}.6H\_\{2\}\\
O solution\$

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  (accurately) weight 200mg of Calcium chloride(?).
\item
  transfer to 100ml volumetric flask, and make up to the mark with
  deionised water.
\end{enumerate}

\hypertarget{task-two}{%
\subsection{task two}\label{task-two}}

0.01M EDTA solution.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  (accurately) weight 200mg of ETDA disodium salt. transfer to 250ml
  volumetric flask, and make up to the mark with deionised water.
\end{enumerate}

\#\#Task A

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Weight out (accurately) 200mg of Zeolite.
\item
  add to 100ml volumetric flask.
\item
  add 75ml of \(CaCl_{2}\) solution.
\item
  Add a stirrer bar. (stir for 5 min)
\item
  Gravity filter the solution 250ml conical flask. (removing zeolite and
  stirrer bar)
\item
  Pipette 25ml of filtered solution into another 250ml conical flask.
\item
  add 3ml of pH 10 buffer.
\item
  Pinch of Eriochrome black T, while swirling.
\item
  Titrate with EDTA solution prepared. (end point colour change to
  blue.)
\end{enumerate}

\hypertarget{part-b}{%
\subsection{Part B}\label{part-b}}

\hypertarget{task-1-concurrent-task-2}{%
\subsubsection{task 1 (concurrent task
2)}\label{task-1-concurrent-task-2}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Add 15ml Glacial acetic acid, 15ml 1-pentanol, and 1ml Concentrated
  sulphuric acid to a 50ml round bottomed flask.
\item
  swirl gentle and add boiling chip
\item
  attach condenser and boil gentle under reflux for 1 hour. (on a
  heating mantle)
\end{enumerate}

\hypertarget{task-2}{%
\subsubsection{task 2}\label{task-2}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Add 15ml Glacial acetic acid, 15ml 1-pentanol, and 2g Zeolite to a
  50ml round bottomed flask.\\
\item
  swirl gentle and add boiling chip
\item
  attach condenser and boil gentle under reflux for 1 hour. (on a
  heating mantle)
\end{enumerate}

\hypertarget{task-3}{%
\subsubsection{task 3}\label{task-3}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  after cooling fiter zeolite mixture
\item
  pour mixture into separating funnel (containing 50ml water)
\item
  stopper funnel and shake gently
\item
  Allow to separate, and run off lower layer (containg water, acetic
  acid and other impurities)
\end{enumerate}

\hypertarget{task-4}{%
\subsubsection{task 4}\label{task-4}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  repeat filtering process with Sulphuric acid solution.
\end{enumerate}

\hypertarget{task-5}{%
\subsubsection{Task 5}\label{task-5}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  wash both solutions with 25ml water
\end{enumerate}

\hypertarget{results}{%
\section{results}\label{results}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Weight of Calcium Chloride(?) for solution 1= \_\_\_\_\_
\item
  Weight of EDTA disodium salt for solution 2= \_\_\_\_\_
\item
  weight of Zeolite.=\_\_\_\_\_\_\_\_\_\_\_\_
\item
  Titration Table
\end{enumerate}

\begin{longtable}[]{@{}llll@{}}
\toprule
& Final Reading & Initial Reading & Volume Added\tabularnewline
\midrule
\endhead
Titration\#1 & & &\tabularnewline
Titration\#2 & & &\tabularnewline
Titration\#3 & & &\tabularnewline
Titration\#4 & & &\tabularnewline
\bottomrule
\end{longtable}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Density and smell of oil, Zeolite solution\_\_\_\_\_\_
\item
  Density and smell of oil, sulphuric acid solution\_\_\_\_\_\_
\end{enumerate}

\hypertarget{questions.}{%
\section{Questions.}\label{questions.}}

\end{document}
