---
title: "Organic Chemistry, Lab 5: carbon-carbon bond formation in aldol and diels-Alder Reactions"
author: "Keelan Krinsky"
output: pdf_document
---

# Submission Information

Name: _Keelan Krinksy_ \newline 
Student number: _1634953_ \newline
Locker number: _365_
Group: _7_ \newline
Experiment Number: _O4_ \newline
Demonstrator: _Thabo_ \newline
Date of Experiment: _02/08/2018_ \newline
Title: _Synthesis and purification, of carboxylic acid, by purification._ \newline

# Reaction Scheme
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 

$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 
$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 

$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 

$\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad$\newline 


# Yield Calculations.

## Part A

```{r, echo=FALSE}
# Table of Reactant details and properties.


# Table Columns
R<-c("Maleic Anhydride","Cyclopentadiene")
M<-c("$0.201g$","$0.786g\\cdot ml^{-1} \\cdot 0.20ml=0.16g$")
N<-c("$\\left(\\frac{0.201g}{98.06g \\cdot mol^{-1}}\\right)= 2.05\\cdot 10^{-3}$","$\\left(\\frac{0.16g}{66.10g \\cdot mol^{-1}}\\right)= 2.4 \\cdot 10^{-1}$")
MW<-c("$98.06$","$66.10$")  

help(exp)

T1<-data.frame("1"=R, "2"=MW, "3"=M, "4"=N)

library(knitr)
colnames(T1)<-c("\\textbf{Reactants Used}", "\\textbf{Molecular Weight($g \\cdot mol^{-1}$)}", "\\textbf{Mass($g$)}","\\textbf{Number of Moles(mol)}")

library("kableExtra")
T1d_latex<-knitr::kable(T1,"latex", digits = 2,row.names=FALSE, col.names = NA, align="llll", caption = "Limiting Reagent Calculations", escape = FALSE) 
kable_styling(T1d_latex, latex_options = c("striped","hold_position"), position= "left")


```
Limiting Reagent: _Maleic Acid_\newline
Molecular Weight of Product: _$164.16g \cdot mol^{-1}$_\newline  
Mass of Product: $8.0149g-7.864g=0.1509g$\newline 
Theoretical Yield: $2.05 \cdot 10^{-3}mol \cdot 164.16g \cdot mol^{-1}=0.337g$\newline
Percentage Yield: $\frac{1.509g}{0.337g}\cdot 100\%=44.8\%$ \newline 



## Part B
```{r, echo=FALSE}
# Table of Reactant details and properties.


# Table Columns
R<-c("Benzyldehyde","Acetone")
M<-c("$1.044g\\cdot ml^{-1} \\cdot 0.60ml=0.63g$","$0.7845g\\cdot ml^{-1} \\cdot 0.22ml=0.17g$")
N<-c("$\\left(\\frac{0.63g}{106.12g \\cdot mol^{-1}}\\right)= 5.9\\cdot 10^{-3}$","$\\left(\\frac{0.17g}{58.08g \\cdot mol^{-1}}\\right)= 2.9 \\cdot 10^{-3}$")
MW<-c("$106.12$","$58.08$")  

help(exp)

T1<-data.frame("1"=R, "2"=MW, "3"=M, "4"=N)

library(knitr)
colnames(T1)<-c("\\textbf{Reactants Used}", "\\textbf{Molecular Weight($g \\cdot mol^{-1}$)}", "\\textbf{Mass($g$)}","\\textbf{Number of Moles(mol)}")

library("kableExtra")
T1d_latex<-knitr::kable(T1,"latex", digits = 2,row.names=FALSE, col.names = NA, align="llll", caption = "Limiting Reagent Calculations", escape = FALSE) 
kable_styling(T1d_latex, latex_options = c("striped","hold_position"), position= "left")


```
Limiting Reagent: _Acetone_\newline
Molecular Weight of Product: _$234.29 \cdot mol^{-1}$_\newline  
Mass of Product: $8.5849g-8.0090g=0.1509g$\newline 
Theoretical Yield: $2.9\cdot 10^{-3}mol \cdot 234.29 g \cdot mol^{-1}=0.68g$\newline
Percentage Yield: $\frac{0.5759g}{0.68g}\cdot 100\%=85\%$ \newline 

